---
title: "REIMB source vocabulary"
output:
  github_document:
    toc: true
    df_print: kable
---

```{r setup, include=FALSE}
rm(list = ls())

knitr::opts_chunk$set(echo = FALSE)

# Document variables 
N_TOPS <- 10

library(tidyverse)
library(lubridate)
library(readxl)

source("../../lib/read_tables.r")

path_vocabulary_tables <- "../../lib/OMOP_VOCABULARY/"

```

## Import Vocabulary
REIMB codes  were downloaded from [dono](dono) (3.4.2020). Missing codes, translations and validity dates were manually colledted from kela's web page into table [./1_source_vocabulary/Kela_reimbursement_codes.xlsx](./1_source_vocabulary/Kela_reimbursement_codes.xlsx) by [\@CoderMikolaj](github.com/CoderMikolaj). 

Resulting table contains the codes, names, and validity periods. Some codes (e.g 307) have a different names during different periods, and therefore codes are repeated on the table for name and time period. 

TODO: we got the name changes over time only for the top 21 codes. The rest should be checked. (top 21 bcs were the ones with no name at all kela's )

Missing English translations were made using Google-tralator


```{r load, message=FALSE, warning=FALSE}
reimb_source <- read_excel("./Kela_reimbursement_codes.xlsx") %>% 
  rename(code = CODE, 
         name_fi = FIN_TXT, 
         name_en = ENG_TXT,
         name_en_source = ENG_SRC ) %>% 
  mutate(name_en_source = str_replace(name_en_source, "GOOGE", "Google")) %>% 
  mutate(code = as.character(code), valid_start_date = dmy(valid_start_date), valid_end_date = dmy(valid_end_date)) 
 
```


## Traslation to English 
The concept names were translated to English as follow: 

- `THL`: concept name in English already exists in the source table. 
- `Traslator`: missing names were send to a translator by <mary.reeve@helsinki.fi>

```{r}
reimb_source %>% count(name_en_source, sort = T)
```

## Formating source vocabulary to OMOP

#### Create VOCABULARY table
```{r}
reimb_vocabulaly <- tibble(
  vocabulary_id = "REIMB", 
  vocabulary_name = "Finnish reimbursement code",
  vocabulary_reference = "", 
  vocabulary_version = "", 
  vocabulary_concept_id = 1
)

reimb_vocabulary_concept <- tibble(
   concept_id = 1,                                                                
   concept_name = "Finnish reimbursement code",
   domain_id  = "Metadata",                                                            
   vocabulary_id  =  "Vocabulary",                                                             
   concept_class_id = "Vocabulary",                                                             
   standard_concept = as.character(NA),                                                                      
   concept_code  =   "OMOP generated",                                                         
   valid_start_date = as.Date("1970-01-01"),                                                                     
   valid_end_date = as.Date("2099-12-31"),                                                                  
   invalid_reason =  as.character(NA)   
)

write_csv(reimb_vocabulaly, "./VOCABULARY_REIMB.csv")

reimb_vocabulaly
```

#### Create CONCEPT_CLASS table
```{r}
reimb_concept_class <- tibble(
  concept_class_id = "REIMB", 
  concept_class_name = "Finnish reimbursement code",
  concept_class_concept_id = 2
) 
 # TOASK: any level or classes ??

reimb_concept_class_concept <- tibble(
   #concept_id = 1,                                                                
   #concept_name = "NOMESCO Finnish modification",
   domain_id  = "Metadata",                                                            
   vocabulary_id  =  "Concept Class",                                                             
   concept_class_id = "Concept Class",                                                             
   standard_concept = as.character(NA),                                                                      
   concept_code  =   "OMOP generated",                                                         
   valid_start_date = as.Date("1970-01-01"),                                                                     
   valid_end_date = as.Date("2099-12-31"),                                                                  
   invalid_reason =  as.character(NA)   
)

reimb_concept_class_concept <- 
  bind_cols(
    reimb_concept_class %>% 
      rename(concept_id = concept_class_concept_id, concept_name = concept_class_id) %>% 
      select(concept_id, concept_name), 
    reimb_concept_class_concept
  )


write_csv(reimb_concept_class, "./CONCEPT_CLASS_REIMB.csv")

reimb_concept_class
```


#### Create CONCEPT table 

At the moment `domain_id` is "Condition" for all although several are not procedures. 

TODO: recalcualte `domain_id` based on the mapping 

```{r}

# add concept_id
reimb_source <- reimb_source %>% 
   mutate( concept_id = 1:nrow(.)+2 ) # as already 2 for domain and concept_class


reimb_concept <- reimb_source %>% 
  transmute(
    concept_id = concept_id, 
    concept_name = name_en,
    domain_id = "Condition", # at the moment
    vocabulary_id = "REIMB", 
    concept_class_id = "REIMB", 
    standard_concept = as.character(NA), 
    concept_code = code, 
    valid_start_date = if_else(is.na(valid_start_date), as.Date("1970-01-01"), valid_start_date), 
    valid_end_date = if_else(is.na(valid_end_date), as.Date("2099-12-31"), valid_end_date), 
    tmp_name_fi = name_fi, # keep for USAGI
    tmp_name_en_source = name_en_source
  ) 
    
        
# add  concepts for domain and concept_class
reimb_concept <- bind_rows(
  reimb_vocabulary_concept, 
  reimb_concept_class_concept, 
  reimb_concept
)


write_csv(reimb_concept, "./CONCEPT_REIMB.csv")

reimb_concept %>% head(N_TOPS)
```


#### Create CONCEPT_SYNONYM table 
- 4181730 # Finnish language
- 4175777 # Swedish language

TOASK: should we add the abbreviation and short_name as synonyms ??


```{r} 
reimb_concept_synomim <-
bind_rows( 
  # Finnish
  reimb_source %>% 
  filter(!is.na(name_fi)) %>% 
  transmute(
    concept_id = concept_id, 
    concept_synonym_name = name_fi, 
    language_concept_id = 4181730
  )
)


write_csv(reimb_concept_synomim, "./CONCEPT_SYNONYM_REIMB.csv")

reimb_concept_synomim %>% head(N_TOPS)

```

Summary of codes added in each language: 

```{r}

reimb_source %>%
  summarise(n_codes = n(), n_witn_name_fi = sum(!is.na(name_fi)))

```




















