---
title: "NOMESCOfi source vocabulary"
output:
  github_document:
    toc: true
    df_print: kable
---

```{r setup, include=FALSE}
rm(list = ls())

knitr::opts_chunk$set(echo = FALSE)

# Document variables 
N_TOPS <- 10

library(tidyverse)
library(lubridate)

source("../../lib/read_tables.r")

path_vocabulary_tables <- "../../lib/OMOP_VOCABULARY/"

```

## Import Vocabulary
NOMESCOfi codes can be found in [Koodistopalvelu](https://91.202.112.142/codeserver/pages/download?name=120_1387444168447.txt&pKey=pubfiles0). 

However, some of the English names are missing and were send to a translator by <mary.reeve@helsinki.fi>. 

Resulting table is [./NOMESCO_codes_UTF8.txt](./NOMESCO_codes_UTF8.txt)

```{r load, message=FALSE, warning=FALSE}
nomescofi_source <- read_delim("./NOMESCO_codes_UTF8.txt", delim = ";", 
                               col_types = cols( .default = col_character() ))
```

```{r}
# NAs added by appeding freqs, remove 
nomescofi_source <- nomescofi_source %>% filter(!is.na(Abbreviation))

# there is one that is repeated WX322 
nomescofi_source <-  nomescofi_source %>%
  filter(!(CodeId=="WX322" & `A:Long_name_source`=="THL")) %>% 
  mutate(BeginningDate  = if_else(CodeId=="WX322", "01/01/2006", BeginningDate ))
```


## Traslation to English 
The concept names were translated to English as follow: 

- `THL`: concept name in English already exists in the source table. 
- `Traslator`: missing names were send to a translator by <mary.reeve@helsinki.fi>

```{r}
nomescofi_source %>% rename(english_name_source = `A:Long_name_source`) %>%  count(english_name_source, sort = T)
```

TODO: the translator and the students notice that few THL translations were wrong, they should be checked 

## Formating source vocabulary to OMOP

#### Create VOCABULARY table
```{r}
nomescofi_vocabulaly <- tibble(
  vocabulary_id = "NOMESCOfi", 
  vocabulary_name = "NOMESCO Finnish modification",
  vocabulary_reference = "https://koodistopalvelu.kanta.fi/codeserver/pages/classification-view-page.xhtml?classificationKey=57&versionKey=119", 
  vocabulary_version = "", 
  vocabulary_concept_id = 1
)

nomescofi_vocabulary_concept <- tibble(
   concept_id = 1,                                                                
   concept_name = "NOMESCO Finnish modification",
   domain_id  = "Metadata",                                                            
   vocabulary_id  =  "Vocabulary",                                                             
   concept_class_id = "Vocabulary",                                                             
   standard_concept = as.character(NA),                                                                      
   concept_code  =   "OMOP generated",                                                         
   valid_start_date = as.Date("1970-01-01"),                                                                     
   valid_end_date = as.Date("2099-12-31"),                                                                  
   invalid_reason =  as.character(NA)   
)

write_csv(nomescofi_vocabulaly, "./VOCABULARY_NOMESCOfi.csv")

nomescofi_vocabulaly
```

#### Create CONCEPT_CLASS table
```{r}
nomescofi_concept_class <- tibble(
  concept_class_id = "NOMESCOfi Level ", 
  concept_class_name = "NOMESCO Finnish Modification Hierarchy Level ",
  concept_class_concept_id = 2
) %>% bind_cols( tibble( level = 0:4)) %>% 
  mutate(
    concept_class_id = str_c(concept_class_id, level ), 
    concept_class_name = str_c(concept_class_name, level ), 
    concept_class_concept_id = concept_class_concept_id + level
    ) %>% 
  select(-level)


nomescofi_concept_class_concept <- tibble(
   #concept_id = 1,                                                                
   #concept_name = "NOMESCO Finnish modification",
   domain_id  = "Metadata",                                                            
   vocabulary_id  =  "Concept Class",                                                             
   concept_class_id = "Concept Class",                                                             
   standard_concept = as.character(NA),                                                                      
   concept_code  =   "OMOP generated",                                                         
   valid_start_date = as.Date("1970-01-01"),                                                                     
   valid_end_date = as.Date("2099-12-31"),                                                                  
   invalid_reason =  as.character(NA)   
)

nomescofi_concept_class_concept <- 
  bind_cols(
    nomescofi_concept_class %>% 
      rename(concept_id = concept_class_concept_id, concept_name = concept_class_id) %>% 
      select(concept_id, concept_name), 
    nomescofi_concept_class_concept
  )


write_csv(nomescofi_concept_class, "./CONCEPT_CLASS_NOMESCOfi.csv")

nomescofi_concept_class
```


#### Create CONCEPT table 

At the moment `domain_id` is "Procedure" for all although several are not procedures. 

TODO: recalcualte `domain_id` based on the mapping 

```{r}

# add concept_id
nomescofi_source <- nomescofi_source %>% 
   mutate( concept_id = 1:nrow(.)+6 ) # as already 6 for domain and concept_class


nomescofi_concept <- nomescofi_source %>% 
  transmute(
    concept_id = concept_id, 
    concept_name = `A:Long_name`,
    domain_id = "Procedure", # at the moment
    vocabulary_id = "NOMESCOfi", 
    concept_class_id = str_c("NOMESCOfi Level ", HierarchyLevel), 
    standard_concept = as.character(NA), 
    concept_code = CodeId, 
    valid_start_date = dmy(BeginningDate), 
    valid_end_date = dmy(ExpiringDate), 
    tmp_name_fi = LongName # keep for USAGI
  ) 
    
        
# add  concepts for domain and concept_class
nomescofi_concept <- bind_rows(
  nomescofi_vocabulary_concept, 
  nomescofi_concept_class_concept, 
  nomescofi_concept
)


write_csv(nomescofi_concept, "./CONCEPT_NOMESCOfi.csv")

nomescofi_concept %>% head(N_TOPS)
```


#### Create CONCEPT_SYNONYM table 
- 4181730 # Finnish language
- 4175777 # Swedish language

TOASK: should we add the abbreviation and short_name as synonyms ??


```{r} 
nomescofi_concept_synomim <-
bind_rows( 
  # Finnish
  nomescofi_source %>% 
  filter(!is.na(LongName)) %>% 
  transmute(
    concept_id = concept_id, 
    concept_synonym_name = LongName, 
    language_concept_id = 4181730
  ),
  # Swedish 
  nomescofi_source %>% 
  filter(!is.na(`A:Långt_namn`)) %>% 
  transmute(
    concept_id = concept_id, 
    concept_synonym_name = `A:Långt_namn`, 
    language_concept_id = 4175777
  )
)


write_csv(nomescofi_concept_synomim, "./CONCEPT_SYNONYM_NOMESCOfi.csv")

nomescofi_concept_synomim %>% head(N_TOPS)

```

Summary of codes added in each language: 

```{r}

nomescofi_source %>%
  summarise(n_codes = n(), n_witn_name_fi = sum(!is.na(LongName)), n_witn_name_se = sum(!is.na(`A:Långt_namn`)))

```




















