---
title: "ICPC"
output:
  github_document:
    toc: true
    df_print: kable
#   html_document:
#     df_print: paged
#     toc: true
---

```{r setup, include=FALSE}
rm(list = ls())

knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(scales)

source("../lib/vocabulary_reporting_fuctions.r")


# Document variables 
N_TOPS <- 10

# output status
db_statuses <- tibble()

```

```{r load, message=FALSE, warning=FALSE}
icpc_concept <- read_csv("./1_source_vocabulary/CONCEPT_ICPC.csv") %>% 
  filter(domain_id != "Metadata")

icpc_mapsto <- read_csv("./2_mapping_to_standard/CONCEPT_RELATIONSHIP_ICPC.csv")

icpc_freqs <- read_csv("./3_freq_of_source_codes/freq_total.csv")
```


# Intro

TODO:




# Formating source vocabulary to OMOP
`ICPC2_merged_and__ICPC1_from_DF6_UTF8.txt`
 Contains the ICPC codes with English and Finnish names.

 This table comes from several sources. Created by  mikolaj.wojnicki@helsinki.fi.
 
```{r}
icpc_concept %>%  count(concept_class_id, sort = T)
```
 

Details in  [./1_source_vocabulary/README.md](./1_source_vocabulary/README.md)


# Mapping the source vocabulary to the standard vocabularies
In short, the source table already includes a column with the equivalent ICD10 codes. This was mapped to the OMOP's ICD10who `concept_id`. 


## Progess in number of codes
```{r}
n_codes <- icpc_concept %>% nrow()
n_codes_accepted <- icpc_mapsto %>% distinct(concept_id_1) %>% nrow()
```


From `r scales::number(n_codes)` codes `r scales::number(n_codes_accepted)` have been approved. 

This makes `r scales::percent(n_codes_accepted/n_codes)` of codes approved. 


```{r}
db_statuses <- bind_rows(
  tibble(
    status = "mapped",
    n_codes = n_codes_accepted,
    per_codes = percent(n_codes_accepted/!!n_codes),
    n_events = as.double(NA),
    per_events = as.character(NA), 
    db_name = "source"
  ),
 tibble(
    status = "not_mapped",
    n_codes = !!n_codes-n_codes_accepted,
    per_codes = percent((!!n_codes-n_codes_accepted)/!!n_codes),
    n_events =  as.double(NA),
    per_events = as.character(NA), 
    db_name = "source"
  )
)
```





# Assessing coverage of databases


```{r  results='asis'}

#write_csv(db_statuses, "status_table.csv" )
 database_coverage_md("ICPC", icpc_concept, icpc_mapsto, icpc_freqs ) 

```
















