---
title: "ICD9fi source vocabulary"
output:
  github_document:
    toc: true
    df_print: kable
---

```{r setup, include=FALSE}
rm(list = ls())

knitr::opts_chunk$set(echo = FALSE)

# Document variables 
N_TOPS <- 10

library(tidyverse)
library(lubridate)
library(readxl)

source("../../lib/read_tables.r")

path_vocabulary_tables <- "../../lib/OMOP_VOCABULARY/"

```

## Import Vocabulary

TODO: i got it from aki now in the gitlab repo 

The concept names are a mixed of Latin, Finnish, and English !!. These are not translated to English !!. The English name were append by matching the ICD9fi code to the ICD9 code dowloaded from [Wolfbane.com](http://www.wolfbane.com/icd/) (1.6.2020). 

```{r}
icd9fi_source <- read_excel("./THL_ICD9fi_matched_ICD9CM_and_Wolfbane_ICD9.xlsx")
```



## Traslation to English 
TODO: so far we have tried to match ICD9fi codes to ICD9CM and Wolfbane (seems to be the same) with  no success 

ATM i will keep the names as they are and lets see later 

## Formating source vocabulary to OMOP

#### Create VOCABULARY table
```{r}
icd9fi_vocabulaly <- tibble(
  vocabulary_id = "ICD9fi", 
  vocabulary_name = "ICD9 Finnish modification",
  vocabulary_reference = "", 
  vocabulary_version = "", 
  vocabulary_concept_id = 1
)

icd9fi_vocabulary_concept <- tibble(
   concept_id = 1,                                                                
   concept_name = "ICD9 Finnish modification",
   domain_id  = "Metadata",                                                            
   vocabulary_id  =  "Vocabulary",                                                             
   concept_class_id = "Vocabulary",                                                             
   standard_concept = as.character(NA),                                                                      
   concept_code  =   "OMOP generated",                                                         
   valid_start_date = as.Date("1970-01-01"),                                                                     
   valid_end_date = as.Date("2099-12-31"),                                                                  
   invalid_reason =  as.character(NA)   
)

write_csv(icd9fi_vocabulaly, "./VOCABULARY_ICD9fi.csv")

icd9fi_vocabulaly
```

#### Create CONCEPT_CLASS table
```{r}
icd9fi_concept_class <- tibble(
  concept_class_id = "-char ICD9fi", 
  concept_class_name = " ICD9 Finnish Modification Hierarchy Level ",
  concept_class_concept_id = 2
) %>% bind_cols( tibble( level = 3:5)) %>% 
  mutate(
    concept_class_id = str_c(level, concept_class_id ), 
    concept_class_name = str_c(level, concept_class_name ), 
    concept_class_concept_id = concept_class_concept_id + level
    ) %>% 
  select(-level)


icd9fi_concept_class_concept <- tibble(
   #concept_id = 1,                                                                
   #concept_name = "NOMESCO Finnish modification",
   domain_id  = "Metadata",                                                            
   vocabulary_id  =  "Concept Class",                                                             
   concept_class_id = "Concept Class",                                                             
   standard_concept = as.character(NA),                                                                      
   concept_code  =   "OMOP generated",                                                         
   valid_start_date = as.Date("1970-01-01"),                                                                     
   valid_end_date = as.Date("2099-12-31"),                                                                  
   invalid_reason =  as.character(NA)   
)

icd9fi_concept_class_concept <- 
  bind_cols(
    icd9fi_concept_class %>% 
      rename(concept_id = concept_class_concept_id, concept_name = concept_class_id) %>% 
      select(concept_id, concept_name), 
    icd9fi_concept_class_concept
  )


write_csv(icd9fi_concept_class, "./CONCEPT_CLASS_ICD9fi.csv")

icd9fi_concept_class
```


#### Create CONCEPT table 

At the moment `domain_id` is "Condition" for all although several are not procedures. 

TODO: recalcualte `domain_id` based on the mapping 

```{r}

# add concept_id
icd9fi_source <- icd9fi_source %>% 
   mutate( concept_id = 1:nrow(.)+7 ) # as already 6 for domain and concept_class


icd9fi_concept <- icd9fi_source %>% 
  transmute(
    concept_id = concept_id, 
    concept_name = ICD9TXT,
    domain_id = "Condition", # at the moment
    vocabulary_id = "ICD9fi", 
    concept_class_id = str_c(nchar(ICD9), "-char ICD9fi"), 
    standard_concept = as.character(NA), 
    concept_code = ICD9, 
    valid_start_date = as.Date("1970-01-01"), 
    valid_end_date = as.Date("2099-12-31"), 
    tmp_name_fi = concept_name  # keep for USAGI
  ) 
    
        
# add  concepts for domain and concept_class
icd9fi_concept <- bind_rows(
  icd9fi_vocabulary_concept, 
  icd9fi_concept_class_concept, 
  icd9fi_concept
)


write_csv(icd9fi_concept, "./CONCEPT_ICD9fi.csv")

icd9fi_concept %>% head(N_TOPS)
```










