---
title: "ICPC source vocabulary"
output:
  github_document:
    toc: true
    df_print: kable
---

```{r setup, include=FALSE}
rm(list = ls())

knitr::opts_chunk$set(echo = FALSE)

# Document variables 
N_TOPS <- 10

library(tidyverse)
library(lubridate)

source("../../lib/read_tables.r")

path_vocabulary_tables <- "../../lib/OMOP_VOCABULARY/"

```

## Import Vocabulary

`ICPC2_merged_and__ICPC1_from_DF6_UTF8.txt`
 Contains the ICPC codes with English and Finnish names.

 This table comes from the ontology gitlab repo. Created by  [\@CoderMikolaj](github.com/CoderMikolaj).

 It combines
 - ICPD2 download from [kodistopalvelu](https://version.helsinki.fi/ontology-group/ontologists/-/wikis/uploads/ce2ce4426b394ef58021b198b28869d4/Koodistopalvelu_ICPC2_v5_2020-04-03.xlsx)  03.04.2020 `standard_ICPC2_kodistopalvelu_282_1471432260621.txt`
 - ICPD2_procedures download from [kodistopalvelu](https://version.helsinki.fi/ontology-group/ontologists/-/wikis/uploads/ca4885e34d5715029f3305e4fb222960/Koodistopalvelu_ICPC2_prosessikoodit_v5_2020-04-03.xlsx)  03.04.2020 `standard_ICPC2_procedures_kodistopalvelu_283_1472198856551.txt`
 - English names downloaded from [ Norwegian Directorate of eHealth](https://ehelse.no/kodeverk/icpc-2e--english-version) `icpc-2e-v7.0.xlsx`.
 - ICPC1 downloaded from ??


```{r load, message=FALSE, warning=FALSE}
icpc_source <- read_tsv("./ICPC2_merged_and__ICPC1_from_DF6_UTF8.txt")

icpc_source <- icpc_source %>%  
  rename(code = Id_FIN, name_fi=`Long name_FIN`, name_en = shortTitle_ENG, main_icd10fi = `Ensisijainen ICD-10_FIN` ) %>% 
  mutate(ICPC_version = if_else(str_detect(code, "^-"), "ICPC2proc", str_c("ICPC", ICPC_version)))
```


## Traslation to English 
All codes have english name 

## Formating source vocabulary to OMOP

#### Create VOCABULARY table
```{r}
icpc_vocabulaly <- tibble(
  vocabulary_id = "ICPC", 
  vocabulary_name = "ICPC1, ICPC2, ICPC2proc",
  vocabulary_reference = "", 
  vocabulary_version = "", 
  vocabulary_concept_id = 1
)

icpc_vocabulary_concept <- tibble(
   concept_id = 1,                                                                
   concept_name = "ICPC1, ICPC2, ICPC2proc",
   domain_id  = "Metadata",                                                            
   vocabulary_id  =  "Vocabulary",                                                             
   concept_class_id = "Vocabulary",                                                             
   standard_concept = as.character(NA),                                                                      
   concept_code  =   "OMOP generated",                                                         
   valid_start_date = as.Date("1970-01-01"),                                                                     
   valid_end_date = as.Date("2099-12-31"),                                                                  
   invalid_reason =  as.character(NA)   
)

write_csv(icpc_vocabulaly, "./VOCABULARY_ICPC.csv")

icpc_vocabulaly
```

#### Create CONCEPT_CLASS table
```{r}
icpc_concept_class <-
bind_rows(
  tibble(
    concept_class_id = "ICPC1", 
    concept_class_name = "ICPC version 1",
    concept_class_concept_id = 2
  ),
  tibble(
    concept_class_id = "ICPC2", 
    concept_class_name = "ICPC version 2",
    concept_class_concept_id = 3
  ),
  tibble(
    concept_class_id = "ICPC2proc", 
    concept_class_name = "ICPC procedures",
    concept_class_concept_id = 4
  )
)
  
  
icpc_concept_class_concept <- tibble(
   #concept_id = 1,                                                                
   #concept_name = "ICPC Finnish modification",
   domain_id  = "Metadata",                                                            
   vocabulary_id  =  "Concept Class",                                                             
   concept_class_id = "Concept Class",                                                             
   standard_concept = as.character(NA),                                                                      
   concept_code  =   "OMOP generated",                                                         
   valid_start_date = as.Date("1970-01-01"),                                                                     
   valid_end_date = as.Date("2099-12-31"),                                                                  
   invalid_reason =  as.character(NA)   
)

icpc_concept_class_concept <- 
  bind_cols(
    icpc_concept_class %>% 
      rename(concept_id = concept_class_concept_id, concept_name = concept_class_id) %>% 
      select(concept_id, concept_name), 
    icpc_concept_class_concept
  )


write_csv(icpc_concept_class, "./CONCEPT_CLASS_ICPC.csv")

icpc_concept_class
```


#### Create CONCEPT table 

At the moment `domain_id` is "Condition" for all although several are not procedures. 

TODO: recalcualte `domain_id` based on the mapping 

```{r}

# add concept_id
icpc_source <- icpc_source %>% 
   mutate( concept_id = 1:nrow(.)+4 ) # as already 6 for domain and concept_class


icpc_concept <- icpc_source %>% 
  transmute(
    concept_id = concept_id, 
    concept_name = name_en,
    domain_id = if_else(ICPC_version == "ICPC2proc", "Procedure", "Condition"), # at the moment
    vocabulary_id = "ICPC", 
    concept_class_id = ICPC_version, 
    standard_concept = as.character(NA), 
    concept_code = code, 
    valid_start_date = as.Date("1970-01-01"),                                                                     
    valid_end_date = as.Date("2099-12-31"),   
    tmp_main_icd10fi  = main_icd10fi, # keep for mapping, 
    tmp_name_fi = name_fi, 
    tmp_name_en_source = "THL"
  ) 
    
        
# add  concepts for domain and concept_class
icpc_concept <- bind_rows(
  icpc_vocabulary_concept, 
  icpc_concept_class_concept, 
  icpc_concept
)


write_csv(icpc_concept, "./CONCEPT_ICPC.csv")

icpc_concept %>% head(N_TOPS)
```


#### Create CONCEPT_SYNONYM table 
- 4181730 # Finnish language
- 4175777 # Swedish language

TOASK: should we add the abbreviation and short_name as synonyms ??


```{r} 
icpc_concept_synomim <-
bind_rows( 
  # Finnish
  icpc_source %>% 
  filter(!is.na(name_fi)) %>% 
  transmute(
    concept_id = concept_id, 
    concept_synonym_name = name_fi, 
    language_concept_id = 4181730
  )
)


write_csv(icpc_concept_synomim, "./CONCEPT_SYNONYM_ICPC.csv")

icpc_concept_synomim %>% head(N_TOPS)

```

Summary of codes added in each language: 

```{r}

icpc_source %>%
  summarise(n_codes = n(), n_witn_name_fi = sum(!is.na( name_en)))

```




















