---
title: "Auto mapping ICD10fi to ICD10who"
output: 
    github_document:
        df_print: tibble
        toc: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
rm(list = ls())

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(scales)
library(stringdist)
```


# Load ICD10fi and ICD10who 

```{r load, message=FALSE}

col_types = cols(
  CodeId1=col_character(), 
  CodeId2=col_character()
)

ICD10fi <- read_csv("../1_source_vocabulary/ICD10_for_source_vocabulary_extended.csv", col_types = col_types)


ICD10who_OMOP <- read_csv("./OMOP_ICD10who.csv") %>% select(concept_code, concept_name)


```

# Study codes

Total number of codes
```{r}
ICD10fi  %>% nrow
```

In the Finnish version of the ICD10 there are two types of additional codes.

- Classification codes: describe ranges of codes other than the conventional ICD10 hierarchy
    - `CodeA-CodeB` : from CodeA to CodeB
- Reason codes: combine codes to add more info on what caused the diagnose, there are 4 marks
   - `CodeA*CodeB` : ”Oirekoodi”, - -CodeA etiology/due to CodeB --
   - `CodeA+CodeB` : ”Syykoodi”, -- CodeB happening in  CodeA--
   - `CodeA#CodeB` : ATC-koodi, CodeB is and ATC code indicating the medicine that caused CodeA
   - `CodeA&CodeB` : ”Kasvainkoodi”, CodeB is and endocrinological disorder code that caused CodeA

Several codes cover the same concept. They have the same LongName and ShortName. 

```{r}
# same_concept <- ICD10fi %>% count(name_fi, sort = T) %>% filter(n>1) %>%  .$name_fi
# 
# ICD10fi %>% filter(name_fi %in% same_concept) %>% arrange(name_fi) %>% 
#   select(code, name_en, code1, code2)# %>%  print(n=100)

```

This is can be due to two reasons: 
- A concept exists on the table as to be append to others "C55", and to be used as diagnosis "c55&". 
- Same concept described as ”Oirekoodi” and ”Syykoodi”, e.g. J17.3*B77.8 (Pneumonia in ascariasis) and B77.8+J17.3 (ascariasis caused by Pneumonia)


**Distribution**
```{r}
ICD10fi  %>% count(code_class)
```

## Matchind codes to ICD10who

Try to match icd10fi's "1 code" codes to icd10who by code and name. 

Also try to match icd10fi's "2 code" to icd10who by name. 

### How many "1 code" are a direct match ??

If matching directly to ICD10fi code
```{r}
ICD10fi_1code <- ICD10fi  %>% filter(code_class=="1 code")

ICD10fi_1code_full_match <- 
  inner_join(ICD10fi_1code,
             ICD10who_OMOP %>% rename(code = concept_code),
             by = "code")


ICD10fi_1code_full_match %>% nrow
```

If matching to ICD10fi code without any union-marks 
```{r}
ICD10fi_1code <- ICD10fi  %>% filter(code_class=="1 code")

ICD10fi_1code_full_match <- 
  inner_join(ICD10fi_1code,
             ICD10who_OMOP %>% rename(code1 = concept_code),
             by = "code1")


ICD10fi_1code_full_match %>% nrow
```

```{r}
ICD10fi_1code_no_match <- ICD10fi  %>% filter(code_class=="1 code") %>% filter(!(code %in% ICD10fi_1code_full_match$code))
```



### How well these agree in the definition ??

Count the differences in characters (string distance) between the English names of icd10fi and icd10who full matches

```{r}
ICD10fi_1code_full_match <- ICD10fi_1code_full_match  %>% 
        mutate(English_name_match_level = stringdist(name_en,concept_name) ) 

ICD10fi_1code_full_match %>% 
        count(English_name_match_level) %>% head(20)
```


Most (8993) match the name exactly!!, 226 NA for these with no English name

```{r}
ICD10fi_1code_almost_match   <- ICD10fi_1code_full_match %>% filter(English_name_match_level > 0)
```


#### Expert checked almost matches 
```{r}
ICD10fi_1code_almost_match %>% 
  select(code, name_en, concept_name, English_name_match_level) %>% 
  arrange(English_name_match_level) %>% 
  write_csv("icd10fi_almost_match_icd10who.csv")
```

These almost matching are saved in [icd10fi_almost_match_icd10who.csv](icd10fi_almost_match_icd10who.csv). 

Name matching was confirmed or rejected by had by Helmi and save as [icd10fi_almost_match_icd10who_CHECKED.csv](icd10fi_almost_match_icd10who_CHECKED.csv).

```{r message=FALSE}
ICD10fi_1code_almost_match_CHECKED <- read_csv("icd10fi_almost_match_icd10who_CHECKED.csv")

not_matching_name_ids <- ICD10fi_1code_almost_match_CHECKED %>%  filter(accept==0) %>% .$CodeId
```

```{r}

ICD10fi_1code_no_match <- bind_rows(
  ICD10fi_1code_no_match, 
  ICD10fi_1code_full_match %>% filter(code %in% not_matching_name_ids)
)

ICD10fi_1code_full_match <- ICD10fi_1code_full_match %>% filter(!(code %in% not_matching_name_ids)) %>% 
  mutate(Mapping_type="ICD10who code and name_en", code_ICD10who = code1)
  

nrow(ICD10fi_1code_full_match)+nrow(ICD10fi_1code_no_match) == nrow(ICD10fi_1code)

```


### How many concepts are in ICD10who but have a different code?
Some concepts are repeated with a different ICD10fi code.
Find these in "2 code" and `ICD10fi_1code_no_match` by having the same name_fi

```{r}

ICD10fi_2code_1code_not_matched <- bind_rows(
  ICD10fi  %>% filter(code_class == "2 code"), 
  ICD10fi_1code_no_match
) 

ICD10fi_2code_1code_name_match <- 
  inner_join(
    ICD10fi_2code_1code_not_matched,
    ICD10fi_1code_full_match  %>% select(code_ICD10who, name_fi), 
    by = c("name_fi")
  ) %>% 
  mutate(Mapping_type="ICD10who name_fi", code_ICD10who = code1)

ICD10fi_2code_1code_name_match %>%  select(code, code_ICD10who, name_fi, name_en, name_en_source) %>% arrange(name_fi)
```

```{r}

ICD10fi_2code_1code_not_matched <- ICD10fi_2code_1code_not_matched %>% filter(!(code %in% ICD10fi_2code_1code_name_match$code))

```

## Prepare un matched "1 code" and "2 code" for USAGI


```{r}
icd10fi_freqs <- read_csv("../3_freq_of_source_codes/freq_total.csv")


# join  -------------------------------------------------------------------------
icd10fi_source_for_usagi <- left_join(
  ICD10fi_2code_1code_not_matched %>% mutate(code1 = str_replace(code1,"\\.", ""), code2 = str_replace(code2,"\\.", "")),
  icd10fi_freqs,
  by=c("code1", "code2")
) 

icd10fi_source_for_usagi <- icd10fi_source_for_usagi %>%
  mutate_if(is.numeric, ~if_else(is.na(.), rep(0, length(.)), .)) %>% 
  arrange(-freq_total)


#check not duplicates introduced
nrow(icd10fi_source_for_usagi)==nrow(ICD10fi_2code_1code_not_matched)

# save --------------------------------------------------------------------
write_csv(icd10fi_source_for_usagi, "./USAGI/icd10fi_source_for_usagi.csv")

```
Append after usagi

```{r}
col_types = cols( .default = col_character(), 
                  sourceFrequency = col_integer(),
                  matchScore  = col_double(),
                  conceptId  = col_integer()
                  )

icd10fi_notmatched_after_usagi <- read_csv("./USAGI/icd10fi_after_usagi.csv", col_types = col_types)
```

```{r}
ICD10fi_2code_1code_not_matched_after_usagi <- left_join(
  ICD10fi_2code_1code_not_matched, 
  icd10fi_notmatched_after_usagi %>% rename(code=sourceCode, Mapping_type=mappingStatus) %>%  select(code, Mapping_type, sourceFrequency  ) %>% distinct(),
  by="code"
) %>% 
  mutate(Mapping_type = if_else(is.na(Mapping_type), "UNCHECKED", Mapping_type))
```


join 1 code 2 code
```{r}
ICD10fi_2code_1code_processed <- bind_rows(
  ICD10fi_1code_full_match,
  ICD10fi_2code_1code_name_match,
  ICD10fi_2code_1code_not_matched_after_usagi
) %>%  
  select(-concept_name, -English_name_match_level)

#check consitecy
nrow(ICD10fi_2code_1code_processed) - nrow(ICD10fi %>% filter(code_class %in% c("1 code", "2 code")))
```

# Check mapping status in "2 code generated"
```{r}
ICD10fi_2codegenerated <- ICD10fi %>% filter(code_class=="2 code generated")
```


```{r}
# join code1 info
ICD10fi_2codegenerated_tmp1 <- left_join(
  ICD10fi_2codegenerated, 
  ICD10fi_2code_1code_processed %>% select(code, Mapping_type) %>% rename(CodeId1=code, Mapping_type_code1=Mapping_type), 
  by="CodeId1"
)

ICD10fi_2codegenerated_tmp2 <- left_join(
  ICD10fi_2codegenerated_tmp1, 
  ICD10fi_2code_1code_processed %>% select(code, Mapping_type) %>% rename(CodeId2=code, Mapping_type_code2=Mapping_type), 
  by="CodeId2"
)

```

```{r}
ICD10fi_2codegenerated<- ICD10fi_2codegenerated_tmp2 %>% 
  mutate(Mapping_type = 
           if_else( Mapping_type_code1=="UNCHECKED" | Mapping_type_code2=="UNCHECKED",
                    "missing one",
                    "mappeable"
                    )
    ) %>% 
  select(-Mapping_type_code1, -Mapping_type_code2)
```

```{r}
ICD10fi_2codegenerated %>% count(Mapping_type)
```





# join all a save
```{r}
icd10_mapped <- bind_rows(
  ICD10fi %>% filter(code_class=="classification"),
  ICD10fi_2code_1code_processed,
  ICD10fi_2codegenerated
)
```
```{r}
icd10_mapped %>% count(code_class)
```

```{r}
write_csv(icd10_mapped, "./icd10_extension_mapped.csv")
```







