---
title: "ICD10fi mappings"
output: 
    github_document:
        df_print: kable
---

```{r setup, include=FALSE}
rm(list = ls())

knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

library(tidyverse)
library(scales)
library(stringdist)
source("../../lib/read_tables.r")

path_vocabulary_tables <- "../../lib/OMOP_VOCABULARY/"

# Document variables 
N_TOPS <- 10
```

```{r message=FALSE}
icd10fi_concept <- read_csv("../1_source_vocabulary/CONCEPT_ICD10fi.csv", 
  col_types = cols(
    concept_id = col_double(),
    concept_name = col_character(),
    domain_id = col_character(),
    vocabulary_id = col_character(),
    concept_class_id = col_character(),
    standard_concept = col_character(),
    concept_code = col_character(),
    valid_start_date = col_date(format = ""),
    valid_end_date = col_date(format = ""),
    invalid_reason = col_character(),
    tmp_code1_concept_id = col_double(),
    tmp_code2_concept_id = col_double(),
    tmp_name_fi = col_character(),
    tmp_code_class = col_character(),
    tmp_code1_code = col_character(),
    tmp_code2_code = col_character()
  )                        
) %>% 
  filter(domain_id != "Metadata")

icd10fi_freqs <- read_csv("../3_freq_of_source_codes/freq_total.csv")
```

## NOTES 

Several codes cover the same concept. They have the same LongName and ShortName. 

```{r}
# same_concept <- icd10fi_concept %>% count(concept_name, sort = T) %>% filter(n>1) %>%  .$name_fi
# 
# ICD10fi %>% filter(name_fi %in% same_concept) %>% arrange(name_fi) %>% 
#   select(code, name_en, code1, code2)# %>%  print(n=100)

```


This is can be due to two reasons: 
- A concept exists on the table as to be append to others "C55", and to be used as diagnosis "c55&". 
- Same concept described as ”Oirekoodi” and ”Syykoodi”, e.g. J17.3*B77.8 (Pneumonia in ascariasis) and B77.8+J17.3 (ascariasis caused by Pneumonia)

**Distribution**
```{r}
icd10fi_concept  %>% count(tmp_code_class)
```

## 1. Mapping type : ICD10who + USAGUI 

First, we map codes to ICD10who by code and name

Second, codes that didn't find a map to ICD10who are mapped to standard-OMOP using USAGI 

### 1.1 Matchind codes to ICD10who
```{r}
icd10who_concept <- read_table_concept(path_vocabulary_tables) %>% 
  filter(vocabulary_id == "ICD10")
```


Try to match icd10fi's "1 code" codes to icd10who by code and name. 

Also try to match icd10fi's "2 code" to icd10who by name. 


##### How many "1 code" are a direct match  to ICD10who??

If matching directly to ICD10fi code
```{r}
ICD10fi_1code <- icd10fi_concept  %>% filter(tmp_code_class=="1 code")

ICD10fi_1code_full_match <- 
  inner_join(ICD10fi_1code,
             icd10who_concept %>% rename_all(str_c, "_who") %>%  rename(concept_code=concept_code_who),
             by = "concept_code")


ICD10fi_1code_full_match %>% nrow
```

If matching to ICD10fi code without any union-marks 
```{r}
ICD10fi_1code <- icd10fi_concept  %>% filter(tmp_code_class=="1 code")

ICD10fi_1code_full_match <- 
  inner_join(ICD10fi_1code  %>% mutate(concept_code_0 = str_replace(concept_code, "\\+|\\*|\\#|\\&", "" )),
             icd10who_concept %>% rename_all(str_c, "_who") %>%  mutate(concept_code_0=concept_code_who),
             by = "concept_code_0") %>% 
  select(-concept_code_0)


ICD10fi_1code_full_match %>% nrow
```
```{r}
ICD10fi_1code_no_match <- ICD10fi_1code %>% 
  filter(!(concept_code %in% ICD10fi_1code_full_match$concept_code ))
```


##### How well these agree in the definition ??

Count the differences in characters (string distance) between the English names of icd10fi and icd10who full matches

```{r}
ICD10fi_1code_full_match <- ICD10fi_1code_full_match  %>% 
        mutate(English_name_match_level = stringdist(concept_name,concept_name_who) ) 

ICD10fi_1code_full_match %>% 
        count(English_name_match_level) %>% head(20)
```

Most (8993) match the name exactly!!, 

```{r}
ICD10fi_1code_almost_match   <- ICD10fi_1code_full_match %>% filter(English_name_match_level > 0)
```


##### Expert checked almost matches 
```{r}
ICD10fi_1code_almost_match %>% 
  select(concept_code, concept_name, concept_name_who, English_name_match_level) %>% 
  arrange(English_name_match_level) %>% 
  write_csv("icd10fi_almost_match_icd10who.csv")
```

These almost matching are saved in [icd10fi_almost_match_icd10who.csv](icd10fi_almost_match_icd10who.csv). 

Name matching was confirmed or rejected by had by Helmi and save as [icd10fi_almost_match_icd10who_CHECKED.csv](icd10fi_almost_match_icd10who_CHECKED.csv).

```{r message=FALSE}
ICD10fi_1code_almost_match_CHECKED <- read_csv("icd10fi_almost_match_icd10who_CHECKED.csv")

not_matching_name_ids <- ICD10fi_1code_almost_match_CHECKED %>%  filter(accept==0) %>% .$CodeId 
```

```{r}

ICD10fi_1code_no_match <- bind_rows(
  ICD10fi_1code_no_match, 
  ICD10fi_1code_full_match %>% filter(concept_code %in% not_matching_name_ids)
)

ICD10fi_1code_full_match <- ICD10fi_1code_full_match %>% filter(!(concept_code %in% not_matching_name_ids)) %>% 
  mutate(tmp_mapping_type="ICD10who code and name_en", concept_code_who = concept_code)
  

nrow(ICD10fi_1code_full_match)+nrow(ICD10fi_1code_no_match) == nrow(ICD10fi_1code)

```


##### How many concepts are in ICD10who but have a different code?
Some concepts are repeated with a different ICD10fi code.
Find these in "2 code" and `ICD10fi_1code_no_match` by having the same tmp_name_fi(icd10fi)

```{r}

ICD10fi_2code_1code_not_matched <- bind_rows(
  icd10fi_concept  %>% filter(tmp_code_class == "2 code"), 
  ICD10fi_1code_no_match
) 

ICD10fi_2code_1code_name_match <- 
  inner_join(
    ICD10fi_2code_1code_not_matched %>% select(-ends_with("_who")),
    ICD10fi_1code_full_match  %>% select(tmp_name_fi, ends_with("_who")), 
    by = c("tmp_name_fi")
  ) %>% 
  mutate(tmp_mapping_type="ICD10who name_fi")

ICD10fi_2code_1code_name_match %>%  select(concept_code, concept_code_who, tmp_name_fi, concept_name) %>% arrange(tmp_name_fi) %>%  head(N_TOPS)
```

```{r}

ICD10fi_2code_1code_not_matched <- ICD10fi_2code_1code_not_matched %>% 
  filter(!(concept_id %in% ICD10fi_2code_1code_name_match$concept_id))

```

```{r}

ICD10fi_mapped_icd10who <- bind_rows(
  ICD10fi_1code_full_match,
  ICD10fi_2code_1code_name_match
)

```


### Create CONCEPT_RELATIONSHIP  for  ICD10who matcc

```{r}
concept_relationship_mapsto <- read_table_concept_relationship_mapsto(path_vocabulary_tables)
```


```{r}

icd10fi_mapsto_icd10who <- left_join(
  ICD10fi_mapped_icd10who %>% select(concept_id, concept_id_who, tmp_mapping_type, tmp_code_class, domain_id_who) %>% rename(concept_id_1 = concept_id, tmp_domain_id_2 = domain_id_who), 
  concept_relationship_mapsto %>% rename(concept_id_who = concept_id_1), 
  by="concept_id_who"
)


icd10fi_mapsto_icd10who %>%  head(N_TOPS)
```


### 1.2 Matchind unmatched "1 code" and "2 code" fwith USAGUI
##### Prepare for USAGI

```{r}
# join  -------------------------------------------------------------------------
icd10fi_source_for_usagi <- left_join(
  ICD10fi_2code_1code_not_matched %>%
    mutate(code1 = str_replace(tmp_code1_code,"\\.", ""), code2 = str_replace(tmp_code1_code,"\\.", "")),
  icd10fi_freqs,
  by=c("code1", "code2")
) 

icd10fi_source_for_usagi <- icd10fi_source_for_usagi %>%
  select(concept_id, concept_code, concept_name, tmp_name_fi, freq_tays, freq_finngen, freq_total ) %>% 
  mutate_if(is.numeric, ~if_else(is.na(.), rep(0, length(.)), .)) %>% 
  arrange(-freq_total)


#check not duplicates introduced
nrow(icd10fi_source_for_usagi)==nrow(ICD10fi_2code_1code_not_matched)

# save --------------------------------------------------------------------
write_csv(icd10fi_source_for_usagi, "./USAGI/icd10fi_source_for_usagi.csv")

```
##### After USAGI

```{r}
col_types = cols( .default = col_character(), 
                  sourceFrequency = col_integer(),
                  matchScore  = col_double(),
                  conceptId  = col_integer()
                  )

icd10fi_notmatched_after_usagi <- read_csv("./USAGI/icd10fi_after_usagi.csv", col_types = col_types)
```

```{r}
ICD10fi_2code_1code_not_matched_after_usagi <- left_join(
  ICD10fi_2code_1code_not_matched, 
  icd10fi_notmatched_after_usagi %>% rename(concept_code=sourceCode, tmp_mapping_type=mappingStatus) %>% 
    select(concept_code, tmp_mapping_type, sourceFrequency, conceptId ),
  by="concept_code"
) %>% 
  mutate(tmp_mapping_type = if_else(is.na(tmp_mapping_type), "UNCHECKED", tmp_mapping_type))
```


##### Create CONCEPT_RELATIONSHIP wfrom  USAGI
```{r}
# load concept table
concept <- read_table_concept(path_vocabulary_tables)
```

```{r}
# join to mapped concepts 
icd10fi_mapping_usagui <- left_join(
  ICD10fi_2code_1code_not_matched_after_usagi %>% 
    filter(tmp_mapping_type == "APPROVED") %>% 
    select(-ends_with("who"), -English_name_match_level ) %>% 
    rename(concept_id_1 = concept_id, concept_id_2 = conceptId), 
  concept %>% 
    select(concept_id, valid_start_date, valid_end_date, domain_id, concept_name) %>% 
    rename_all(str_c, "_2"), 
  by="concept_id_2"
)

# TOFIX : removing nonsensical domains 
icd10fi_mapping_usagui <- icd10fi_mapping_usagui %>% 
  filter(domain_id_2 %in% c("Condition", "Observation", "Procedure"))


# select max valid_start_date and min for valid_end_date 
icd10fi_mapsto_usagi <- icd10fi_mapping_usagui %>% 
  transmute(
    concept_id_1 = concept_id_1,
    concept_id_2 = concept_id_2,
    relationship_id = "Maps to",
    valid_start_date = pmax(valid_start_date, valid_start_date_2), 
    valid_end_date   = pmin(valid_end_date, valid_end_date_2), 
    invalid_reason = as.character(NA), 
    tmp_mapping_type = "USAGI", 
    tmp_code_class = tmp_code_class, 
    tmp_domain_id_2 = domain_id_2
  )


icd10fi_mapsto_usagi %>%  head(N_TOPS)

```

### 2 Create CONCEPT_RELATIONSHIP  for 2 code generated 

```{r}
icd10fi_mapsto_icd10who_and_usagi <- bind_rows(
  icd10fi_mapsto_icd10who,
  icd10fi_mapsto_usagi
) %>% select(-concept_id_who)

```


```{r}
icd10fi_2codegenerated <- icd10fi_concept %>% filter(tmp_code_class=="2 code generated") %>% 
  select(concept_id, tmp_code1_concept_id, tmp_code2_concept_id, valid_start_date, valid_end_date) %>% 
  gather("source", "concept_id_compound", -concept_id, -valid_start_date, -valid_end_date) %>%
  select(-source)
```


```{r}
# join code1 info
icd10fi_2codegenerated_mapping <- left_join(
  icd10fi_2codegenerated, 
  icd10fi_mapsto_icd10who_and_usagi %>% rename(concept_id_compound = concept_id_1 ), 
  by="concept_id_compound"
)

icd10fi_2codegenerated_mapsto <- icd10fi_2codegenerated_mapping %>% 
  transmute(
    concept_id_1 = concept_id, 
    concept_id_2 = concept_id_2, 
    relationship_id = relationship_id, 
    valid_start_date = pmax(valid_start_date.x, valid_start_date.y ), 
    valid_end_date = pmin(valid_end_date.x, valid_end_date.y ), 
    invalid_reason = invalid_reason, 
    tmp_mapping_type = tmp_mapping_type, 
    tmp_code_class ="2 code generated", 
    tmp_domain_id_2 = tmp_domain_id_2
  ) %>% 
  distinct()


# find concept_id_1s that are missing a mapping in concept_id_2
concept_id_1s_with_concept_id_2_missing <- icd10fi_2codegenerated_mapsto %>% 
  filter(is.na(tmp_mapping_type)) %>% .$concept_id_1

icd10fi_2codegenerated_mapsto <- icd10fi_2codegenerated_mapsto %>% 
  filter(!is.na(tmp_mapping_type)) %>% 
  mutate(tmp_mapping_type = 
           if_else(concept_id_1 %in% concept_id_1s_with_concept_id_2_missing, 
                    "missing one",
                    "fully mapped"
                 )
         )  
```




```{r}
icd10_mapsto <- bind_rows(
  icd10fi_mapsto_icd10who_and_usagi, 
  icd10fi_2codegenerated_mapsto
) %>% select(1, 4:8, tmp_mapping_type, tmp_code_class, tmp_domain_id_2)


write_csv(icd10_mapsto, "./CONCEPT_RELATIONSHIP_ICD10fi.csv")


icd10_mapsto %>% head(N_TOPS)
```



## Recalculate domain 
```{r}
icd10fi_domain <- icd10_mapsto %>% 
  rename(concept_id = concept_id_1) %>% 
  group_by(concept_id) %>% 
  summarise(domain = str_c( sort(unique(tmp_domain_id_2))  , collapse = " "), .groups="drop") %>% 
  mutate(domain = str_replace(domain, "Condition Observation", "Condition/Obs" )) %>% 
  mutate(domain = str_replace(domain, "Condition Procedure", "Condition/Procedure" ))
  
```

```{r}
icd10fi_concept_new <- left_join(
  icd10fi_concept, 
  icd10fi_domain, 
  by="concept_id"
) %>% 
  mutate(domain_id = if_else(is.na(domain), domain_id, domain ))
```




















