---
title: "ICD9fi mappings"
output: 
    github_document:
        df_print: kable
---

```{r setup, include=FALSE}
rm(list = ls())

knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

library(tidyverse)
source("../../lib/read_tables.r")

path_vocabulary_tables <- "../../lib/OMOP_VOCABULARY/"

# Document variables 
N_TOPS <- 10
```

## Mapping type : USAGUI 
#### Preparation fotrUSAGI

In short, USAGI directly on icd9fi's source names (Latin, Finnish, or English) ranked by frequency calculated from FinnGen-DF5.

The frequencies from FinnGen-DF5 [3_freq_of_source_codes/freq_total.csv](3_freq_of_source_codes/freq_total.csv)
were append to  [1_source_vocabulary/CONCEPT_ICD9fi.csv](1_source_vocabulary/CONCEPT_ICD9fi.csv). 


Mapping was carried by medical student [\@kalleaseppala](github.com/kalleaseppala). 
 


```{r message=FALSE}
icd9fi_concept <- read_csv("../1_source_vocabulary/CONCEPT_ICD9fi.csv") %>%  
  filter(vocabulary_id =="ICD9fi")
icd9fi_freqs <- read_csv("../3_freq_of_source_codes/freq_total.csv")
```

```{r message=FALSE}
icd9fi_source_for_usagi <- left_join(
  icd9fi_concept,
  icd9fi_freqs %>% rename(concept_name = code),
  by="concept_name"
) 
```

```{r}
icd9fi_source_for_usagi <- icd9fi_source_for_usagi %>%
  mutate_if(is.numeric, ~if_else(is.na(.), rep(0, length(.)), .)) %>% 
  arrange(-freq_total)

write_csv(icd9fi_source_for_usagi, "./USAGI/icd9fi_source_for_usagi.csv")
```


#### After USAGI

```{r}
icd9fi_after_usagi <- read_csv("./USAGI/icd9fi_after_usagi.csv", 
  col_types = cols( 
    .default = col_character(), 
    sourceFrequency = col_integer(),
    matchScore  = col_double(),
    conceptId  = col_integer()
  )
) %>%  # cant get the encoding right %>%
  filter(mappingStatus!="0.3793565332889557") # not sure what it is 
```
Summary of mapping progress after USAGI:

```{r}
icd9fi_after_usagi %>% distinct(sourceCode, mappingStatus) %>% count(mappingStatus)
```

```{r}
a <- icd9fi_after_usagi %>% distinct(sourceCode, mappingStatus, sourceFrequency)

n_total_envents <- sum(a$sourceFrequency)
n_approved_events <- a %>% filter(mappingStatus=="APPROVED") %>% .$sourceFrequency %>% sum()

max_n_events_unmapped <-  a %>% filter(mappingStatus=="UNCHECKED") %>% .$sourceFrequency %>%  max()
```

Accepted codes covers `r scales::percent(n_approved_events/n_total_envents)` of the total number of events in the combined databases. 

Accepted codes covers all codes with more than `r scales::number(max_n_events_unmapped)` events in the combined databases.

Top`r N_TOPS` of the unchecked events sort by number of events : 

```{r}
icd9fi_after_usagi %>% filter(mappingStatus=="UNCHECKED") %>% arrange(-sourceFrequency) %>% select(sourceCode,  sourceFrequency, matchScore, sourceName, `ADD_INFO:possible_english_name`) %>%  head(N_TOPS)
```

## Formating mappigs to OMOP

Mappings based on USAGI. 

- For valid_start_date max of source concept and mapped concept is chosen 
- For valid_end_date min of source concept and mapped concept is chosen 

```{r}
# load concept table
concept <- read_table_concept(path_vocabulary_tables)
```


```{r}
# join to icd9fi concepts
icd9fi_mapping <- inner_join(
  icd9fi_concept %>% 
    select(concept_id, concept_code, valid_start_date, valid_end_date) %>% 
    rename(concept_id_1 = concept_id),
  icd9fi_after_usagi %>%  filter(mappingStatus == "APPROVED") %>% 
    select(sourceCode, conceptId) %>% 
    rename(concept_id_2 = conceptId, concept_code = sourceCode),
  by="concept_code"
) %>% select(-concept_code )



# join to mapped concepts 
icd9fi_mapping <- left_join(
  icd9fi_mapping, 
  concept %>% 
    select(concept_id, valid_start_date, valid_end_date, domain_id) %>% 
    rename_all(str_c, "_2"), 
  by="concept_id_2"
)

# TOFIX : removing nonsensical domains 
icd9fi_mapping <- icd9fi_mapping %>% 
  filter(domain_id_2 %in% c("Condition", "Observation", "Procedure"))


# select max valid_start_date and min for valid_end_date 
icd9fi_mapsto <- icd9fi_mapping %>% 
  transmute(
    concept_id_1 = concept_id_1,
    concept_id_2 = concept_id_2,
    relationship_id = "Maps to",
    valid_start_date = pmax(valid_start_date, valid_start_date_2), 
    valid_end_date   = pmin(valid_end_date, valid_end_date_2), 
    invalid_reason = as.character(NA), 
    tmp_mapping_type = "USAGI", 
    tmp_domain_id_2 = domain_id_2
  )


write_csv(icd9fi_mapsto, "CONCEPT_RELATIONSHIP_icd9fi.csv")

icd9fi_mapsto %>%  head(N_TOPS)

```



## Recalculate domain 
```{r}
icd9fi_domain <- icd9fi_mapsto %>% 
  rename(concept_id = concept_id_1) %>% 
  group_by(concept_id) %>% 
  summarise(domain = str_c( sort(unique(tmp_domain_id_2))  , collapse = " "), .groups="drop") %>% 
  mutate(domain = str_replace(domain, "Condition Observation", "Condition/Obs" )) %>% 
  mutate(domain = str_replace(domain, "Condition Procedure", "Condition/Procedure" ))
  
```

```{r message=FALSE, warning=FALSE}
icd9fi_concept <- read_csv("../1_source_vocabulary/CONCEPT_ICD9fi.csv")

icd9fi_concept_new <- left_join(
  icd9fi_concept, 
  icd9fi_domain, 
  by="concept_id"
) %>% 
  mutate(domain_id = if_else(is.na(domain), domain_id, domain )) %>%
  select(-domain)

#overwrite
write_csv(icd9fi_concept_new, "../1_source_vocabulary/CONCEPT_ICD9fi.csv")

```

```{r}
icd9fi_concept_new %>% count(domain_id, sort=T)
```



























