---
title: "Prepare REIMB codes for USAGI"
output: 
    github_document:
        df_print: tibble
---

```{r setup, include=FALSE}
rm(list = ls())

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(scales)
library(readxl)

```

## Load

```{r load, message=FALSE, warning=FALSE}

reimb_standard_eng <- read_excel("./KELA_reimburesement_ENG_translations_Tuomo(1).xlsx") %>% filter(!is.na(Code))
reimb_standard <- read_tsv("./KELA_korvausnumerot_KELAn_sivuilta_kopioitu_2020-04-03.txt")

reimb_standard_eng <- reimb_standard_eng %>% rename(code = Code, name_english = `Name in English`, name_finnish = `Name in Finnish`)

reimb_standard <- reimb_standard %>% 
  mutate(asterisk = str_detect(Korvausnumero, "\\*"), Korvausnumero = as.integer(str_replace(Korvausnumero,"\\*","" ))) %>% 
  rename(code = Korvausnumero, name_finnish = Sairaus) %>% 
  filter(!is.na(code))

reimb <- left_join(reimb_standard, 
                   reimb_standard_eng %>% select(-name_finnish), 
                   by="code") %>% 
  select(code, asterisk, name_finnish, name_english)



freq_reimb_finngen <- read_csv("../1_source_files/freq_reimb_finngen.csv") %>% select(-X1) %>% 
  rename(freq_finngen=n_codes)
```


# Study finnGen

Number of codes in each source: 
```{r}
reimb %>%  nrow()
freq_reimb_finngen %>%  nrow()


n_events_finngen <- freq_reimb_finngen$freq_finngen %>% sum()
```

```{r}
 finngen_join <- full_join(freq_reimb_finngen %>% rename(code = CODE1), 
                           reimb,
                           by = c("code"))
```



**How many codes labeled as icd10fi in FinnGen are not in the icd10fi standard ?**
```{r paged.print=TRUE}
no_finngen_standard <- finngen_join %>% filter(is.na(asterisk)) %>% select(code, freq_finngen ) %>%
  mutate( freq_finngen_per=percent(freq_finngen/n_events_finngen))

no_finngen_standard
```






```{r}
finngen_join <- finngen_join %>%
  mutate( freq_finngen_per=percent(freq_finngen/n_events_finngen))%>%
  mutate(valid_start_date = as.Date(NA), valid_end_date = as.Date(NA), invalid_reason=as.character(NA)) %>% 
  select(code, freq_finngen, freq_finngen_per, 3:9)

finngen_join %>% arrange(-is.na(asterisk)) 


write_csv(finngen_join %>% arrange(-is.na(asterisk)), "./reimb_finngen_freq.csv")
```

This table is based on `KELA_korvausnumerot_KELAn_sivuilta_kopioitu_2020-04-03.txt`.
Reimbursement "code"; "freq_finngen" number of occurrences in DF5; "freq_finngen_per" calculated as the percetage of total ; "n_patients"  number of patients with that code in DF5 ; if the code had an asterisk then "asterisk" is TRUE; "name_finnish";     "name_english" append from `KELA_reimburesement_ENG_translations_Tuomo.xlsx` if there; I added the two columns   "valid_start_date", "valid_end_date", "invalid_reason". 
 
 
 The table is shorted out to show on top the codes that appear in DF5 but dont exist on `KELA_korvausnumerot_KELAn_sivuilta_kopioitu_2020-04-03.txt` (and therefore have no Finnish name). 
 
 TODO: 
 1- Find the name of the missing codes. Explain shortly why they got invalid, and what is the new valid code in "invalid_reason".
 With my limited Finnish I found some info here https://www.kela.fi/2010_laakekorvaukset-17122010. But I'm sure it has to be a better place with this info.  
 2- Find the "valid_start_date" and  "valid_end_date" for all the codes. If there is not end valid bcs it is currently in use you can give "31.12.2099"
 
 
 



