---
title: "ICD8fi source vocabulary"
output:
  github_document:
    toc: true
    df_print: kable
---

```{r setup, include=FALSE}
rm(list = ls())

knitr::opts_chunk$set(echo = FALSE)

# Document variables 
N_TOPS <- 10

library(tidyverse)
library(lubridate)

source("../../lib/read_tables.r")

path_vocabulary_tables <- "../../lib/OMOP_VOCABULARY/"

```

## Import Vocabulary

TODO ask  [\@CoderMikolaj](github.com/CoderMikolaj) how table was created. 

```{r}
icd8fi_source <- read_tsv("./ICD8Merged_Version5_UTF8.txt", 
  col_types = cols(
    Code = col_character(),
    DF6_Freq = col_double(),
    Latin = col_character(),
    Finnish = col_character(),
    Swedish = col_character(),
    English = col_character(),
    Uncertain_language = col_character(),
    Needs_LAT_trans = col_double(),
    Finnish_requires_review = col_double(),
    Needs_FIN_trans = col_double(),
    English_requires_review = col_double(),
    needs_ENG_trans = col_double(),
    Comments = col_character()
  )
) %>% filter(!is.na(Code))
```

Some seem to be classification codes xxx-xxx. Remove these

```{r}
icd8fi_source <- icd8fi_source %>%
  filter(nchar(Code)<7)
```




## Traslation to English 
The concept names were translated to English as follow: 

- `THL`: concept name in English already exists in the source table. 
- `Google`: Missing English translations were made using Google-translator (TODO:)


```{r}
# TODO: temp fix ############################
icd8fi_source <- icd8fi_source %>% 
  filter(!( is.na(Finnish) & is.na(English))) %>%  # remove if we dont have a name eng or fi
  mutate(English = if_else(!is.na(English), English, Finnish)) # if not english set the fi name, will be traslated
#########################################



icd8fi_source %>% 
  mutate(name_en_source = if_else(needs_ENG_trans==0, "THL", "Google")) %>%
  count(name_en_source, sort = T)
```

## Formating source vocabulary to OMOP

#### Create VOCABULARY table
```{r}
icd8fi_vocabulaly <- tibble(
  vocabulary_id = "ICD8fi", 
  vocabulary_name = "ICD8 Finnish modification",
  vocabulary_reference = "", 
  vocabulary_version = "", 
  vocabulary_concept_id = 1
)

icd8fi_vocabulary_concept <- tibble(
   concept_id = 1,                                                                
   concept_name = "ICD8 Finnish modification",
   domain_id  = "Metadata",                                                            
   vocabulary_id  =  "Vocabulary",                                                             
   concept_class_id = "Vocabulary",                                                             
   standard_concept = as.character(NA),                                                                      
   concept_code  =   "OMOP generated",                                                         
   valid_start_date = as.Date("1970-01-01"),                                                                     
   valid_end_date = as.Date("2099-12-31"),                                                                  
   invalid_reason =  as.character(NA)   
)

write_csv(icd8fi_vocabulaly, "./VOCABULARY_ICD8fi.csv")

icd8fi_vocabulaly
```

#### Create CONCEPT_CLASS table
```{r}
icd8fi_concept_class <- tibble(
  concept_class_id = "-digit ICD8fi", 
  concept_class_name = " digits ICD8 Finnish Modification",
  concept_class_concept_id = 2
) %>% bind_cols( tibble( level = 3:5)) %>% 
  mutate(
    concept_class_id = str_c(level, concept_class_id ), 
    concept_class_name = str_c(level, concept_class_name ), 
    concept_class_concept_id = concept_class_concept_id + level
    ) %>% 
  select(-level)


icd8fi_concept_class_concept <- tibble(
   #concept_id = 1,                                                                
   #concept_name = "NOMESCO Finnish modification",
   domain_id  = "Metadata",                                                            
   vocabulary_id  =  "Concept Class",                                                             
   concept_class_id = "Concept Class",                                                             
   standard_concept = as.character(NA),                                                                      
   concept_code  =   "OMOP generated",                                                         
   valid_start_date = as.Date("1970-01-01"),                                                                     
   valid_end_date = as.Date("2099-12-31"),                                                                  
   invalid_reason =  as.character(NA)   
)

icd8fi_concept_class_concept <- 
  bind_cols(
    icd8fi_concept_class %>% 
      rename(concept_id = concept_class_concept_id, concept_name = concept_class_id) %>% 
      select(concept_id, concept_name), 
    icd8fi_concept_class_concept
  )


write_csv(icd8fi_concept_class, "./CONCEPT_CLASS_ICD8fi.csv")

icd8fi_concept_class
```


#### Create CONCEPT table 

At the moment `domain_id` is "Condition" for all although several are not procedures. 

TODO: recalcualte `domain_id` based on the mapping 

```{r}

# add concept_id
icd8fi_source <- icd8fi_source %>% 
   mutate( concept_id = 1:nrow(.)+7 ) # as already 6 for domain and concept_class

icd8fi_concept <- icd8fi_source %>% 
  transmute(
    concept_id = concept_id, 
    concept_name = English,
    domain_id = "Condition", # at the moment
    vocabulary_id = "ICD8fi", 
    concept_class_id = str_c(nchar(Code), "-digit ICD8fi"), 
    standard_concept = as.character(NA), 
    concept_code = Code, 
    valid_start_date = dmy(01011970), 
    valid_end_date = dmy(3112299), 
    tmp_name_fi = Finnish, # keep for USAGI
    tmp_name_en_source = if_else(needs_ENG_trans==0, "THL", "Google")
  ) 
    
        
# add  concepts for domain and concept_class
icd8fi_concept <- bind_rows(
  icd8fi_vocabulary_concept, 
  icd8fi_concept_class_concept, 
  icd8fi_concept
)


write_csv(icd8fi_concept, "./CONCEPT_ICD8fi.csv")

icd8fi_concept %>% head(N_TOPS)
```



















