---
title: "ICD8fi"
output:
  github_document:
    toc: true
    df_print: kable
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
rm(list = ls())

knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(scales)

source("../lib/vocabulary_reporting_fuctions.r")


# Document variables 
N_TOPS <- 10

# output status
db_statuses <- tibble()

```

```{r load, message=FALSE, warning=FALSE}
icd8fi_concept <- read_csv("./1_source_vocabulary/CONCEPT_ICD8fi.csv")

icd8fi_mapsto <- read_csv("../ATC/2_mapping_to_standard/CONCEPT_RELATIONSHIP_ATC.csv") %>% 
  filter(FALSE) ## read an other table to keep same colums

icd8fi_freqs <- read_csv("./3_freq_of_source_codes/freq_total.csv")
```


# Intro
TODO:


# Formating source vocabulary to OMOP
TODO ask  [\@CoderMikolaj](github.com/CoderMikolaj) how table was created. 

TODO: Missing English translations were made using Google-tralator
 

## Progess in number of codes
```{r}
n_codes <- icd8fi_concept %>% nrow()
n_codes_accepted <- icd8fi_mapsto %>% distinct(concept_id_1) %>% nrow()
```


From `r scales::number(n_codes)` codes `r scales::number(n_codes_accepted)` have been approved. 

This makes `r scales::percent(n_codes_accepted/n_codes)` of codes approved. 


```{r}
db_statuses <- bind_rows(
  tibble(
    status = "mapped",
    n_codes = n_codes_accepted,
    per_codes = percent(n_codes_accepted/!!n_codes),
    n_events = as.double(NA),
    per_events = as.character(NA), 
    db_name = "source"
  ),
 tibble(
    status = "not_mapped",
    n_codes = !!n_codes-n_codes_accepted,
    per_codes = percent((!!n_codes-n_codes_accepted)/!!n_codes),
    n_events =  as.double(NA),
    per_events = as.character(NA), 
    db_name = "source"
  )
)
```





# Assessing coverage of databases


```{r  results='asis'}

#write_csv(db_statuses, "status_table.csv" )
 database_coverage_md("ICD8fi", icd8fi_concept, icd8fi_mapsto, icd8fi_freqs) 

```
















