---
title: "add derived codes"
output: 
    github_document:
        df_print: tibble
        toc: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
rm(list = ls())

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(scales)
library(stringdist)
```


# Load ICD10fi

```{r load, message=FALSE}
ICD10fi <- read_delim("./58_1471433699473.txt", ";", 
                               locale = locale(encoding = 'ISO-8859-1'),
                               col_types = cols( .default = col_character() )
                               )

```


# Parse descritions into regex rules 
Get codes with a described expansion in `ALONG:Huom.`

```{r}
icd10fi_expansion_description <- ICD10fi %>% 
  filter(!is.na(`A:Yhdistelymerkki`)) %>% 
  filter(!str_detect(CodeId,"[:digit:]$")) %>% #dont end with mark
  filter(str_detect(`ALONG:Huom.`, "[:upper:][:digit:][:digit:]")) %>% 
  filter(str_detect(`ALONG:Huom.`, "tiologi")) %>% 
  arrange(`ALONG:Huom.`) %>% 
  select(CodeId,`ALONG:Huom.`) 

icd10fi_expansion_description
```

There are 2 types of expansion: 
- Added single code "Etiologiseksi koodiksi valitaan Xxx.x"
- To chose from a chapter/group "Etiologinen koodi valitaan ryhmistä"


## Expand one code addition
Parse description in Huom to CodeId2
```{r}
#code
icd10fi_extended_by_one_code <- 
icd10fi_expansion_description %>% filter(str_detect(`ALONG:Huom.`, "koodik") & !str_detect(`ALONG:Huom.`, "Lisäksi")) %>% 
 mutate(group_text = str_extract(`ALONG:Huom.`, "[:upper:][:digit:][:digit:].*")) %>% 
  mutate(group_text = str_replace(group_text, "(.*)\\s-\\s.*", "\\1")) %>% 
  # tai to , 
  mutate_at(vars(group_text), str_replace_all,"\\stai\\s", "|") %>% 
  #convert to regex
  mutate(group_reg = str_c("^", str_replace_all(group_text,"\\|", "|^")))
```

## Expand by class/chapter/group
```{r}
range_to_regex <- function(match){
  range <- str_extract_all(match, "[:digit:][:digit:]") %>% unlist() %>% as.integer()
  match <- range[1]:range[2] %>% as.character() %>% if_else(nchar(.)==1, str_c("0",.),.) %>% str_c(collapse = "|")
  return(str_c("(",match,")"))
}

#group
icd10fi_extended_by_group <- 
icd10fi_expansion_description %>% filter(str_detect(`ALONG:Huom.`, "luokasta|luvusta|ryhmistä|ryhmästä|rymistä")) %>% 
  #extract codes
  mutate(group_text = str_extract(`ALONG:Huom.`, "[:upper:][:digit:][:digit:].*")) %>% 
  mutate(group_text = str_replace(group_text, "(.*)\\s-\\s.*", "\\1")) %>% 
  mutate(group_text = str_replace(group_text, "(.*)\\.\\s.*", "\\1")) %>% 
  mutate(group_text = str_replace(group_text, "(.*)\\.\\s.*", "\\1")) %>% 
  #extract negations
  mutate(group_text_ei = str_extract(group_text, "\\(ei.*\\)")) %>% 
  mutate(group_text = str_remove(group_text, "\\s\\(ei.*\\)")) %>% 
  mutate(group_text_ei = str_replace(group_text_ei, "\\(ei\\s(.*)\\)", "\\1")) %>% 
  # tai to , 
  mutate_at(vars(group_text, group_text_ei), str_replace_all, "\\stai\\sryhmistä\\s|\\stai\\sryhmästä\\s|,\\s|\\stai\\s", "|") %>%
  # format as regex
  mutate(group_reg=group_text, group_reg_ei=group_text_ei) %>% 
  mutate(group_reg = str_replace(group_reg, "C00-D", "C00-C99\\|D00-D")) %>% 
  mutate(group_reg = str_replace(group_reg, "A([:digit:][:digit:])-B", "A\\1-A99\\|B00-B")) %>% 
  mutate_at(vars(group_reg,group_reg_ei), str_replace_all, "-[:upper:]", "-" ) %>% 
  #mutate_at(vars(group_reg,group_reg_ei), str_replace_all, "([:digit:][:digit:]-[:digit:][:digit:])", "\\[\\1\\]" ) %>% 
  mutate_at(vars(group_reg,group_reg_ei), str_replace_all, "\\|", "|^" ) %>% 
  mutate_at(vars(group_reg,group_reg_ei), ~{str_c("^",.)} ) %>% 
  mutate(group_reg = str_replace_all(group_reg, "[:digit:][:digit:]-[:digit:][:digit:]", range_to_regex)) %>% 
  mutate(group_reg_ei = str_replace_all(group_reg_ei, "50-53", "(50|51|52|53)"))%>% 
  #extract merkki
  mutate(group_reg_4merkki = str_extract(`ALONG:Huom.`, "(\\.[:digit:])($|\\.$)")) %>% 
  mutate(group_reg_4merkki = str_replace(group_reg_4merkki, "\\.$", "")) %>% 
  mutate(group_reg_4merkki = str_c("\\", group_reg_4merkki) ) 

  
  
  #icd10fi_extended_by_group %>% select(group_text, group_reg, group_reg_ei, group_reg_4merkki) %>% print(n=300)

```

Expand rules for  `H36.00*	Etiologinen koodi valitaan tässä ryhmässä ryhmistä E10-E14. Neljäs merkki on .3`

```{r}
a<-ICD10fi %>% filter(ParentId == "H36*") %>% filter(is.na(`A:Koodi2`)) %>%  select(CodeId)
b <- icd10fi_extended_by_group %>% filter(CodeId == "H36.00*")
b$CodeId[1] = a
b <- b %>%  unnest(CodeId)

icd10fi_extended_by_group <- bind_rows(icd10fi_extended_by_group, b) %>%  distinct()

```

```{r}
# join rules 
icd10fi_extended_rules <- bind_rows(icd10fi_extended_by_one_code, icd10fi_extended_by_group)

```


# apply regex rules 

Extend leaves only 
```{r}
icd10fi_leaves <- ICD10fi %>%
  filter(`A:Lehtisolmu`=="T" & is.na(`A:Koodi2`)) %>% 
  select(CodeId)
  
```

apply rules 
```{r}
icd10fi_extended<- icd10fi_extended_rules %>% #slice(1)%>% 
  mutate(code2 = pmap(list(group_reg, group_reg_ei, group_reg_4merkki), 
                      function(group_reg, group_reg_ei, group_reg_4merkki){
                        icd10fi_leaves %>% select(CodeId) %>%  
                          filter(str_detect(CodeId, group_reg)) %>% 
                          filter(is.na(group_reg_ei)      | !str_detect(CodeId, group_reg_ei)) %>% 
                          filter(is.na(group_reg_4merkki) |  str_detect(CodeId, group_reg_4merkki)) %>% 
                          rename(CodeId2 = CodeId)
  })) %>% unnest(code2) 
  
icd10fi_extended <- icd10fi_extended %>% 
  rename(CodeId1 = CodeId) %>% 
  mutate(code1 = str_replace(CodeId1, "\\+|\\*|\\#|\\&", "")) %>% 
  mutate(code2 = str_replace(CodeId2,"\\+|\\*|\\#|\\&", "")) %>% 
  mutate(CodeId = str_c(CodeId1, code2)) %>% 
  select(CodeId, CodeId1, CodeId2, code1, code2) %>%  
  distinct(CodeId,.keep_all = T)
  
icd10fi_extended
#icd10fi_extended_by_group %>%filter( str_detect(CodeId, "N08.3")) %>% view

```


How many already exists ?
```{r}
new_icd10fi_extended <- left_join(icd10fi_extended, 
          ICD10fi %>% select(CodeId) %>%  mutate(is=TRUE), 
          by="CodeId"
) %>%
  filter(is.na(is))

nrow(icd10fi_extended)-nrow(new_icd10fi_extended)
```



## Append the new codes to the standard
```{r}
# get one code source codes 
icd10fi_source <- read_csv("./icd10fi_for_source_vocabulary0.csv")

# TODO: this should be in the format_ICD10 file
icd10fi_source <- left_join(
  icd10fi_source, 
  ICD10fi %>% rename(code=CodeId, code1=`A:Koodi1`, code2=`A:Koodi2`, connector_mark=`A:Yhdistelymerkki`) %>% 
    select(code, code1, code2, connector_mark), 
  by="code"
  ) %>% 
  mutate(code_class = case_when(
    is.na(connector_mark) & !is.na(code2) ~ "classification", 
    !is.na(code2) ~ "2 code", 
    TRUE ~ "1 code"
  ))

icd10fi_source_1_code <- icd10fi_source %>% filter(code_class=="1 code") %>% select(-code1, -code2, - code_class, -connector_mark)
```


```{r}
# join code1 info
icd10_extended_tmp1 <- left_join(
  new_icd10fi_extended, 
  icd10fi_source_1_code %>% rename_all(~str_c("code1_",.)) %>% rename(CodeId1=code1_code), 
  by="CodeId1"
)

# join code2 info
icd10_extended_tmp2 <- left_join(
  icd10_extended_tmp1, 
  icd10fi_source_1_code %>% rename_all(~str_c("code2_",.)) %>% rename(CodeId2=code2_code), 
  by="CodeId2"
)

```


```{r}
#merge cols 
icd10_extended <- icd10_extended_tmp2 %>% mutate(
    #names
    name_fi = str_c(code1_name_fi, " LIITTYVÄ  ",code2_name_fi),
    name_en = str_c(code1_name_en, " RELATED-TO  ",code2_name_en),
    name_se = str_c(code1_name_se, " RELATERAD-TILL  ",code2_name_se),
    name_en_source = case_when(
      code1_name_en_source == "THL"    & code2_name_en_source == "THL"    ~ "THL",
      code1_name_en_source == "THL"    & code2_name_en_source == "Google" ~ "THL+Google",
      code1_name_en_source == "Google" & code2_name_en_source == "THL"    ~ "THL+Google",
      code1_name_en_source == "Google" & code2_name_en_source == "Google" ~ "Google"
    ),
    # dates 
    valid_start_date = max(code1_valid_start_date, code2_valid_start_date),
    valid_end_date   = min(code1_valid_end_date,   code2_valid_end_date),
    #
    code_class = "2 code generated", 
    connector_mark = "*"
  ) %>% 
  rename(code=CodeId) %>% 
  select(code, name_fi, name_en, name_se, valid_start_date, valid_end_date,  name_en_source,  code1,  code2,  connector_mark,  code_class, CodeId1, CodeId2)


#join 
icd10fi <- bind_rows(icd10fi_source, icd10_extended)
```




```{r}
write_csv(icd10fi, "./ICD10_for_source_vocabulary_extended.csv")
```



