---
title: "ICD10fi source vocabulary"
output:
  github_document:
    toc: true
    df_print: kable
---

```{r setup, include=FALSE}
rm(list = ls())

knitr::opts_chunk$set(echo = FALSE)

# Document variables 
N_TOPS <- 10

library(tidyverse)
library(lubridate)
library(scales)

source("../../lib/read_tables.r")

path_vocabulary_tables <- "../../lib/OMOP_VOCABULARY/"

```

## Import Vocabulary
The list of the icd10fi codes have been downloaded from the official source:  [kodistopalvelu](https://91.202.112.142/codeserver/pages/publication-view-page.xhtml?distributionKey=9395&versionKey=58&returnLink=fromVersionPublicationList)(7.5.2020). 

```{r}
icd10fi_source <- read_delim("./58_1471433699473.txt", ";", 
                             locale = locale(encoding = 'ISO-8859-1'),
                             col_types = cols( .default = col_character() )
)
                             
icd10fi_source <- icd10fi_source %>% 
  mutate(code = CodeId, 
         name_fi = LongName, name_en = `A:Long_name`, name_se = `A:Långt_namn`,  
         valid_start_date = BeginningDate, valid_end_date = ExpiringDate,
         code1=`A:Koodi1`, code2=`A:Koodi2`, connector_mark=`A:Yhdistelymerkki`
         ) %>% 
  mutate_at(vars(valid_start_date, valid_end_date), dmy)
```

This table contains all the single codes, some composed-codes, and classification. The last only used for hierarchy and not diagnose. 
In addition, the table also contains the rules to generated other valid composed-codes (in column `A:Huom`).

 
Table describes a total of  `r number(icd10fi_source %>% nrow)` codes distributed as follow: 

```{r}

icd10fi_source <- icd10fi_source %>% 
  mutate(code_class = case_when(
    is.na(connector_mark) & !is.na(code2) ~ "classification", 
    !is.na(code2) ~ "2 code", 
    TRUE ~ "1 code"
  )
)

icd10fi_source %>%  count(code_class, sort = T)
```

We wont be using the classigfication codes 
```{r}
icd10fi_source <- icd10fi_source %>% filter(code_class != "classification")
```



## Traslation to English 
All codes come with the name in Finnish, in Swedish, and many, but not all, also in English. 
For these missing we used Google-translator. 

- `THL`: concept name in English already exists in the source table. 
- `Google`: translated using Google-translator API


```{r}
# Traslate missing using GT

#gl_auth(r"(C:\Users\javier\AppData\Local\R\keys\traslater-e7e9c4dc6948.json)")
#
# icd10fi_source_google_tralations <- icd10fi_source %>%# head(10) %>% 
#   mutate(name_en_source = if_else(!is.na(name_en), "THL",  "Google")) %>% 
#   mutate(name_en = if_else(is.na(name_en), gl_translate(name_fi, source = "fi", target = "en")$translatedText, name_en ))
#   select(code, name_en_source, name_en, name_fi )
#
# write_csv(icd10fi_source_google_tralations, "icd10fi_source_google_tralations.csv" )


icd10fi_source_google_tralations <- read_csv("./icd10fi_source_google_tralations.csv",
  col_types = cols( .default = col_character() )
)


icd10fi_source <- left_join(
  icd10fi_source %>%  select(-name_en, -starts_with("name_fi"),
                             -starts_with("better_name_en"),
                             -starts_with("name_en_source"), 
                             #-starts_with("valid_start_date"),
                             #-starts_with("valid_end_date")
                             ), 
  icd10fi_source_google_tralations %>% select(
    -starts_with("valid_start_date"),
    -starts_with("valid_end_date")
    ),
  by="code"
)

#tmp fix add name_fi
write_csv(
  icd10fi_source  %>% 
    mutate(better_name_en="") %>% 
    select(code, name_en_source, name_fi, name_en, better_name_en), 
  "./icd10fi_source_google_tralations.csv"
)

```



```{r}
icd10fi_source %>% count(name_en_source, sort = T)
```


## Extending composed-codes
We generate new composed-codes based on the "1 code" codes that contain generation rules in column `A:Huom`. 
This process is detailed in [README.Rmd](README.Rmd).

#### Parse descritions into regex rules 
Get codes with a described expansion in `ALONG:Huom.`

```{r}
icd10fi_expansion_description <- icd10fi_source %>% 
  filter(!is.na(`A:Yhdistelymerkki`)) %>% 
  filter(!str_detect(CodeId,"[:digit:]$")) %>% #dont end with mark
  filter(str_detect(`ALONG:Huom.`, "[:upper:][:digit:][:digit:]")) %>% 
  filter(str_detect(`ALONG:Huom.`, "tiologi")) %>% 
  arrange(`ALONG:Huom.`) %>% 
  select(CodeId,`ALONG:Huom.`) 

icd10fi_expansion_description %>% head(N_TOPS)
```

There are 2 types of expansion: 
- Added single code "Etiologiseksi koodiksi valitaan Xxx.x"
- To chose from a chapter/group "Etiologinen koodi valitaan ryhmistä"


#### Expand one code addition
Parse description in Huom to CodeId2
```{r}
#code
icd10fi_extended_by_one_code <- 
icd10fi_expansion_description %>% filter(str_detect(`ALONG:Huom.`, "koodik") & !str_detect(`ALONG:Huom.`, "Lisäksi")) %>% 
 mutate(group_text = str_extract(`ALONG:Huom.`, "[:upper:][:digit:][:digit:].*")) %>% 
  mutate(group_text = str_replace(group_text, "(.*)\\s-\\s.*", "\\1")) %>% 
  # tai to , 
  mutate_at(vars(group_text), str_replace_all,"\\stai\\s", "|") %>% 
  #convert to regex
  mutate(group_reg = str_c("^", str_replace_all(group_text,"\\|", "|^")))
```

#### Expand by class/chapter/group
```{r}
range_to_regex <- function(match){
  range <- str_extract_all(match, "[:digit:][:digit:]") %>% unlist() %>% as.integer()
  match <- range[1]:range[2] %>% as.character() %>% if_else(nchar(.)==1, str_c("0",.),.) %>% str_c(collapse = "|")
  return(str_c("(",match,")"))
}

#group
icd10fi_extended_by_group <- 
icd10fi_expansion_description %>% filter(str_detect(`ALONG:Huom.`, "luokasta|luvusta|ryhmistä|ryhmästä|rymistä")) %>% 
  #extract codes
  mutate(group_text = str_extract(`ALONG:Huom.`, "[:upper:][:digit:][:digit:].*")) %>% 
  mutate(group_text = str_replace(group_text, "(.*)\\s-\\s.*", "\\1")) %>% 
  mutate(group_text = str_replace(group_text, "(.*)\\.\\s.*", "\\1")) %>% 
  mutate(group_text = str_replace(group_text, "(.*)\\.\\s.*", "\\1")) %>% 
  #extract negations
  mutate(group_text_ei = str_extract(group_text, "\\(ei.*\\)")) %>% 
  mutate(group_text = str_remove(group_text, "\\s\\(ei.*\\)")) %>% 
  mutate(group_text_ei = str_replace(group_text_ei, "\\(ei\\s(.*)\\)", "\\1")) %>% 
  # tai to , 
  mutate_at(vars(group_text, group_text_ei), str_replace_all, "\\stai\\sryhmistä\\s|\\stai\\sryhmästä\\s|,\\s|\\stai\\s", "|") %>%
  # format as regex
  mutate(group_reg=group_text, group_reg_ei=group_text_ei) %>% 
  mutate(group_reg = str_replace(group_reg, "C00-D", "C00-C99\\|D00-D")) %>% 
  mutate(group_reg = str_replace(group_reg, "A([:digit:][:digit:])-B", "A\\1-A99\\|B00-B")) %>% 
  mutate_at(vars(group_reg,group_reg_ei), str_replace_all, "-[:upper:]", "-" ) %>% 
  #mutate_at(vars(group_reg,group_reg_ei), str_replace_all, "([:digit:][:digit:]-[:digit:][:digit:])", "\\[\\1\\]" ) %>% 
  mutate_at(vars(group_reg,group_reg_ei), str_replace_all, "\\|", "|^" ) %>% 
  mutate_at(vars(group_reg,group_reg_ei), ~{str_c("^",.)} ) %>% 
  mutate(group_reg = str_replace_all(group_reg, "[:digit:][:digit:]-[:digit:][:digit:]", range_to_regex)) %>% 
  mutate(group_reg_ei = str_replace_all(group_reg_ei, "50-53", "(50|51|52|53)"))%>% 
  #extract merkki
  mutate(group_reg_4merkki = str_extract(`ALONG:Huom.`, "(\\.[:digit:])($|\\.$)")) %>% 
  mutate(group_reg_4merkki = str_replace(group_reg_4merkki, "\\.$", "")) %>% 
  mutate(group_reg_4merkki = str_c("\\", group_reg_4merkki) ) 

  
  
  #icd10fi_extended_by_group %>% select(group_text, group_reg, group_reg_ei, group_reg_4merkki) %>% print(n=300)

```

Expand rules for  `H36.00*	Etiologinen koodi valitaan tässä ryhmässä ryhmistä E10-E14. Neljäs merkki on .3`

```{r}
a<-icd10fi_source %>% filter(ParentId == "H36*") %>% filter(is.na(`A:Koodi2`)) %>%  select(CodeId)
b <- icd10fi_extended_by_group %>% filter(CodeId == "H36.00*")
b$CodeId[1] = a
b <- b %>%  unnest(CodeId)

icd10fi_extended_by_group <- bind_rows(icd10fi_extended_by_group, b) %>%  distinct()

```

```{r}
# join rules 
icd10fi_extended_rules <- bind_rows(icd10fi_extended_by_one_code, icd10fi_extended_by_group)

```


#### apply regex rules 

Extend leaves only 
```{r}
icd10fi_leaves <- icd10fi_source %>%
  filter(`A:Lehtisolmu`=="T" & is.na(`A:Koodi2`)) %>% 
  select(CodeId)
  
```

apply rules 
```{r}
icd10fi_extended<- icd10fi_extended_rules %>% #slice(1)%>% 
  mutate(code2 = pmap(list(group_reg, group_reg_ei, group_reg_4merkki), 
                      function(group_reg, group_reg_ei, group_reg_4merkki){
                        icd10fi_leaves %>% select(CodeId) %>%  
                          filter(str_detect(CodeId, group_reg)) %>% 
                          filter(is.na(group_reg_ei)      | !str_detect(CodeId, group_reg_ei)) %>% 
                          filter(is.na(group_reg_4merkki) |  str_detect(CodeId, group_reg_4merkki)) %>% 
                          rename(CodeId2 = CodeId)
  })) %>% unnest(code2) 
  
icd10fi_extended <- icd10fi_extended %>% 
  rename(CodeId1 = CodeId) %>% 
  mutate(code1 = str_replace(CodeId1, "\\+|\\*|\\#|\\&", "")) %>% 
  mutate(code2 = str_replace(CodeId2,"\\+|\\*|\\#|\\&", "")) %>% 
  mutate(CodeId = str_c(CodeId1, code2)) %>% 
  select(CodeId, CodeId1, CodeId2, code1, code2) %>%  
  distinct(CodeId,.keep_all = T)
  
icd10fi_extended %>% head(N_TOPS)
#icd10fi_extended_by_group %>%filter( str_detect(CodeId, "N08.3")) %>% view

```


How many already exists ?
```{r}
new_icd10fi_extended <- left_join(icd10fi_extended, 
          icd10fi_source %>% select(CodeId) %>%  mutate(is=TRUE), 
          by="CodeId"
) %>%
  filter(is.na(is))

nrow(icd10fi_extended)-nrow(new_icd10fi_extended)
```



#### Append the new codes to the standard
```{r}
# lets set here the concept_id 
icd10fi_source <- icd10fi_source %>% 
  mutate(concept_id = 1:nrow(.))


# get one code source codes 
icd10fi_source_1_code <- icd10fi_source %>% filter(code_class=="1 code") %>%
  select(concept_id, CodeId, name_fi, name_en, name_se, name_en_source, BeginningDate, valid_start_date, valid_end_date)
```


```{r}
# join code1 info
icd10_extended_tmp1 <- left_join(
  new_icd10fi_extended, 
  icd10fi_source_1_code %>% rename_all(~str_c("code1_",.)) %>% rename(CodeId1=code1_CodeId), 
  by="CodeId1"
)

# join code2 info
icd10_extended_tmp2 <- left_join(
  icd10_extended_tmp1, 
  icd10fi_source_1_code %>% rename_all(~str_c("code2_",.)) %>% rename(CodeId2=code2_CodeId), 
  by="CodeId2"
)

```


```{r}
#merge cols 
icd10_extended <- icd10_extended_tmp2 %>% mutate(
    #names
    name_fi = str_c(code1_name_fi, " LIITTYVÄ  ",code2_name_fi),
    name_en = str_c(code1_name_en, " RELATED-TO  ",code2_name_en),
    name_se = str_c(code1_name_se, " RELATERAD-TILL  ",code2_name_se),
    name_en_source = case_when(
      code1_name_en_source == "THL"    & code2_name_en_source == "THL"    ~ "THL",
      code1_name_en_source == "THL"    & code2_name_en_source == "Google" ~ "THL+Google",
      code1_name_en_source == "Google" & code2_name_en_source == "THL"    ~ "THL+Google",
      code1_name_en_source == "Google" & code2_name_en_source == "Google" ~ "Google"
    ),
    # dates 
    valid_start_date = pmax(code1_valid_start_date, code2_valid_start_date),
    valid_end_date   = pmin(code1_valid_end_date,   code2_valid_end_date),
    #
    code_class = "2 code generated", 
    connector_mark = "*", 
    concept_id = (1:nrow(.))+max(icd10fi_source$concept_id)
  ) %>% 
  #rename(code=CodeId) %>% 
  select(concept_id, CodeId, name_fi, name_en, name_se, name_en_source,
         valid_start_date, valid_end_date,   
         code_class, CodeId1, CodeId2, connector_mark, 
         code1_concept_id, code2_concept_id)


#join 
icd10fi_source_extended <- bind_rows(icd10fi_source, icd10_extended)
```

For the new codes, name is given as both source code's names joined by "LIITTYVÄ", "RELATERAD-TILL", and "RELATED-TO" respectively for name_fi, name_se, and name_en. 

For the validity dates, the latest is chose for valid_start_date and the earliest for valid_end_date. 

For the how the English name was translated: `THL` if both were translated by THL, `Google` if both were translated by google-translator, and `THL+Google` if both.  

```{r}
icd10fi_source_extended %>%  count(name_en_source, sort = T)
```



## Formating source vocabulary to OMOP

#### Create VOCABULARY table
```{r}
icd10fi_vocabulaly <- tibble(
  vocabulary_id = "ICD10fi", 
  vocabulary_name = "ICD10fi Finnish modification",
  vocabulary_reference = "https://91.202.112.142/codeserver/pages/publication-view-page.xhtml?distributionKey=9395&versionKey=58&returnLink=fromVersionPublicationList", 
  vocabulary_version = "", 
  vocabulary_concept_id = 1
)

icd10fi_vocabulary_concept <- tibble(
   concept_id = 1,                                                                
   concept_name = "ICD10fi Finnish modification",
   domain_id  = "Metadata",                                                            
   vocabulary_id  =  "Vocabulary",                                                             
   concept_class_id = "Vocabulary",                                                             
   standard_concept = as.character(NA),                                                                      
   concept_code  =   "OMOP generated",                                                         
   valid_start_date = as.Date("1970-01-01"),                                                                     
   valid_end_date = as.Date("2099-12-31"),                                                                  
   invalid_reason =  as.character(NA)   
)

write_csv(icd10fi_vocabulaly, "./VOCABULARY_ICD10fi.csv")

icd10fi_vocabulaly
```

#### Create CONCEPT_CLASS table
```{r}
icd10fi_concept_class <-
  bind_rows(
    tibble(
      concept_class_id = "ICD10fi Hierarchy", # TODO set to the omop code for this 
      concept_class_name = "ICD10fi Hierarchy",
      concept_class_concept_id = 2
    ),
    tibble(
      concept_class_id = "ICD10fi 1 code", 
      concept_class_name = "ICD10fi 1 code",
      concept_class_concept_id = 3
    ),
    tibble(
      concept_class_id = "ICD10fi 2 code", 
      concept_class_name = "ICD10fi Finnish Modification composed code",
      concept_class_concept_id = 5
    ),
    tibble(
      concept_class_id = "ICD10fi 2 code [A]", 
      concept_class_name = "ICD10fi Finnish Modification composed code [auto generated]",
      concept_class_concept_id = 6
    ),
)

icd10fi_concept_class_concept <- tibble(
   #concept_id = 1,                                                                
   #concept_name = "ICD10fi Finnish modification",
   domain_id  = "Metadata",                                                            
   vocabulary_id  =  "Concept Class",                                                             
   concept_class_id = "Concept Class",                                                             
   standard_concept = as.character(NA),                                                                      
   concept_code  =   "OMOP generated",                                                         
   valid_start_date = as.Date("1970-01-01"),                                                                     
   valid_end_date = as.Date("2099-12-31"),                                                                  
   invalid_reason =  as.character(NA)   
)

icd10fi_concept_class_concept <- 
  bind_cols(
    icd10fi_concept_class %>% 
      rename(concept_id = concept_class_concept_id, concept_name = concept_class_id) %>% 
      select(concept_id, concept_name), 
    icd10fi_concept_class_concept
  )


write_csv(icd10fi_concept_class, "./CONCEPT_CLASS_ICD10fi.csv")

icd10fi_concept_class
```


#### Create CONCEPT table 

At the moment `domain_id` is "Condition" for all although several are not procedures. 

TODO: recalculate `domain_id` based on the mapping 

```{r}

# add concept_id
icd10fi_source_extended <- icd10fi_source_extended %>% 
  mutate(concept_id = concept_id+6)


icd10fi_concept <- icd10fi_source_extended %>% 
  transmute(
    concept_id = concept_id, 
    concept_name = name_en,
    domain_id = "Condition", # at the moment
    vocabulary_id = "ICD10fi", 
    concept_class_id = case_when(
      code_class=="2 code generated" ~ "ICD10fi 2 code [A]", 
      code_class=="2 code" ~ "ICD10fi 2 code",
      code_class=="1 code" & HierarchyLevel<=3 ~ "ICD10fi Hierarchy",
      code_class=="1 code" & HierarchyLevel>3 ~ "ICD10fi 1 code"
    ), 
    standard_concept = as.character(NA), 
    concept_code = CodeId, 
    valid_start_date = valid_start_date, 
    valid_end_date = valid_end_date, 
    tmp_code1_concept_id = code1_concept_id, # keep for mapping
    tmp_code2_concept_id = code2_concept_id, # keep for mapping
    tmp_name_fi = name_fi, # keep for USAGI
    tmp_code_class = code_class, 
    tmp_code1_code = `A:Koodi1`, 
    tmp_code2_code = `A:Koodi2`, 
    tmp_name_en_source = name_en_source
  ) 
    
        
# add  concepts for domain and concept_class
icd10fi_concept <- bind_rows(
  icd10fi_vocabulary_concept, 
  icd10fi_concept_class_concept, 
  icd10fi_concept
)


write_csv(icd10fi_concept, "./CONCEPT_ICD10fi.csv")

icd10fi_concept %>% head(N_TOPS)
```


#### Create CONCEPT_SYNONYM table 
- 4181730 # Finnish language
- 4175777 # Swedish language

TOASK: should we add the abbreviation and short_name as synonyms ??


```{r} 
icd10fi_concept_synomim <-
bind_rows( 
  # Finnish
  icd10fi_source_extended %>% 
  filter(!is.na(name_fi)) %>% 
  transmute(
    concept_id = concept_id, 
    concept_synonym_name = name_fi, 
    language_concept_id = 4181730
  ),
  # Swedish 
  icd10fi_source_extended %>% 
  filter(!is.na(name_se)) %>% 
  transmute(
    concept_id = concept_id, 
    concept_synonym_name = name_se, 
    language_concept_id = 4175777
  )
)


write_csv(icd10fi_concept_synomim, "./CONCEPT_SYNONYM_ICD10fi.csv")

icd10fi_concept_synomim %>% head(N_TOPS)

```

Summary of codes added in each language: 

```{r}

icd10fi_source_extended %>%
  summarise(n_codes = n(), n_witn_name_fi = sum(!is.na(name_fi)), n_witn_name_se = sum(!is.na(name_se)))

```




















