---
title: "ATC source vocabulary"
output: 
    github_document:
        df_print: tibble
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
rm(list = ls())

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
source("../../lib/read_tables.r")

path_vocabulary_tables <- "../../lib/OMOP_VOCABULARY/"

```

## Import Vocabulary
ATC is already in OMOP. Can be extract from the vocabularies. 

```{r load, message=FALSE, warning=FALSE}
atc_source <- read_table_concept(path_vocabulary_tables) %>% filter(vocabulary_id == "ATC")
```

## Create CONCEPT table 
```{r}
write_csv(atc_source, "CONCEPT_ATC.csv")
```



## Notes

I notice that few old codes are missing and reported in the forum ([see post](https://forums.ohdsi.org/t/atc-alterations-missing-in-omop/11496/3)). 
In sort, ATC releases do not incude these any more and are out of the OMOP pipeline, but they could be included. 

ATC modifications were downloaded from [https://www.whocc.no/atc_ddd_alterations__cumulative/atc_alterations/](https://www.whocc.no/atc_ddd_alterations__cumulative/atc_alterations/) and shaved in [./atc_modifications.csv](./atc_modifications.csv)

```{r}
atc_changes <- read_csv("./atc_modifications.csv") %>% 
  mutate_at(vars(previous_atc_code, new_atc_code), str_replace, "\\s[:digit:]+\\)", "")
```

```{r}
missing_atc_in_omop <- full_join(
  atc_source %>% select(concept_code, concept_id), 
  atc_changes %>%  rename(concept_code = previous_atc_code),
  by="concept_code"
) %>% 
  filter(is.na(concept_id))
```


```{r}
write_csv(missing_atc_in_omop, "./missing_atc_in_omop.csv")
```

These not in OMOP are saved in [./missing_atc_in_omop.csv](./missing_atc_in_omop.csv)


**In finggen there are also some E codes**

**From Ari Ahola-Olli:** E can be found from there. But the reason why they are not in WHO database is that they are not actual drugs.

